---
title: "graph_merged"
author: "Sasha, Ayliana, Taxi"
date: "July 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### load library
```{r}
library(tidyverse)
library(igraph)
library(lubridate)
library(leaflet)
library(mapview)
```
#### load functions
```{r load functions}
source('../../src/cache.R')
source('../../src/k_shortest.R')
```

#### initializing caching
```{r cache}
init_cache('../../cache/graph_merged')
```

#### Load data
```{r}
cache.read('stops','../../data/google_transit_subway_static/stops.txt')
cache.read('routes_data', '../../data/google_transit_subway_static/routes.txt')
cache.read('trips', '../../data/google_transit_subway_static/trips.txt', col_types = cols(route_id ='c'))
cache.read('stop_times','../../data/google_transit_subway_static/stop_times.txt')
cache.read('transfers', '../../data/google_transit_subway_static/transfers.txt')
```

#### Set filters for time intervals/day
```{r}

# all service
# time_filter <- seq(0, 23)
# day_filter <- c("Weekday", "Saturday", "Sunday")

# ordinary (weekday+daytime) service
time_filter <- seq(6, 23)
day_filter <- c("Weekday")

# weekend (daytime) service
#time_filter <- seq(6, 23)
#day_filter <- c("Saturday", "Sunday")

# late night (weekday+weekend) service
# time_filter <- seq(0, 5)
# day_filter <- c("Weekday", "Saturday", "Sunday")
```

#### Formatting static
```{r}
# all scheduled trips
create_all_trips <- function (){
  stop_times %>%
  filter(!is.na(arrival_time)) %>%
  left_join(stops) %>%
  extract(trip_id, c("day_of_week"), regex=".*-.*-(.*)-.*", remove=FALSE) %>%
  extract(trip_id, c("time"), regex=".*-.*-.*-.*_(.*)_.*\\.\\.?.*", remove=FALSE) %>%
  mutate(direction = substr(stop_id, 4, 4),
         stop_id = substr(stop_id, 1, 3),
         prev_stop_id = ifelse(trip_id == lag(trip_id), lag(stop_id), NA),
         prev_stop_name = ifelse(trip_id == lag(trip_id), lag(stop_name), NA),
         trip_start_time = seconds_to_period(as.numeric(time)*.6),
         trip_start_time = as.POSIXct(sprintf("%s:%s:%s", 
                                              hour(trip_start_time), minute(trip_start_time), second(trip_start_time)),
                                      "%H:%M:%S", tz="America/New_York")) %>%
  left_join(trips) %>%
  select(route_id, trip_id, direction_id = direction, day_of_week, trip_start_time, arrival_time, departure_time, 
         stop_id, stop_name, prev_stop_id, prev_stop_name)
}

cache('all_trips')
```

#### Transfer formatting
```{r}
# prepping the transfer data
create_transfer_sequences <- function(){
  transfers %>% left_join(stops, by = c("to_stop_id" = "stop_id")) %>% 
  left_join(stops, by = c("from_stop_id" = "stop_id")) %>%
  mutate(route_ids = "T", direction_id = "T", weight = min_transfer_time, sd = NA, lower_quartile = NA, median = NA, upper_quartile = NA) %>%
  select(route_ids, direction_id, stop_id = to_stop_id, stop_name = stop_name.x, prev_stop_id = from_stop_id, 
         prev_stop_name = stop_name.y, weight, sd, lower_quartile, median, upper_quartile)
}
cache('transfer_sequences')
```

#### Realtime Formatting
```{r}
create_realtime <- function(){
  load('../../data/todd_subway_realtime.RData')
  realtime %>%
  mutate(day_of_week = weekdays(departure_time),
         day_of_week = ifelse(day_of_week != "Saturday" & day_of_week != "Sunday",
                              "Weekday", day_of_week),
         direction_id = ifelse(direction == 1, "N", "S"))
}
cache('realtime')
```


#### Apply time/day filter here
```{r}
# unique trip sequences by line
filtered_sequences <- all_trips %>%
  filter(hour(trip_start_time) %in% time_filter) %>%
  filter(day_of_week %in% day_filter) 

counts <- filtered_sequences %>%
  group_by(route_id) %>%
  summarize(total = n())

percentages <- filtered_sequences %>%
  select(route_id, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name) %>%
  group_by(route_id, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name) %>%
  summarize(count = n()) %>%
  left_join(counts) %>%
  mutate(relative_percentage = count/total)

quartiles <- percentages %>%
  group_by(route_id) %>%
  summarize(lower_10 = quantile(relative_percentage, 0.1), lower = quantile(relative_percentage, 0.25), 
            lower_30 = quantile(relative_percentage, 0.3),
            median = median(relative_percentage), 
            upper = quantile(relative_percentage, 0.75),
            upper_90 = quantile(relative_percentage, 0.9))

percentages <- percentages %>%
  left_join(quartiles)

unique_sequences <- percentages %>%
  filter(relative_percentage >= lower_10) %>%
  select(route_id, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name) 
```


#### Get weights for the connections between stations
```{r weights, warning=FALSE}
create_station_weights <- function(){
realtime %>%
  mutate(stop_mta_id = substr(stop_mta_id, 1, 3)) %>%
  
  arrange(realtime_trip_id, departure_time) %>% 
  mutate(travel_time = ifelse(realtime_trip_id == lag(realtime_trip_id),
                              departure_time - lag(departure_time), NA)) %>%
  mutate(prev_stop_mta_id = ifelse(realtime_trip_id == lag(realtime_trip_id),
                                   lag(stop_mta_id), NA)) %>%
  filter(!is.na(travel_time), 
         hour(departure_time) %in% time_filter,
         day_of_week %in% day_filter) %>% 
  group_by(stop_mta_id, prev_stop_mta_id) %>% 
  summarize(weight = mean(travel_time), sd = sd(travel_time, na.rm=TRUE),
            lower_quartile = quantile(travel_time, 0.25),
            median = median(travel_time), upper_quartile = quantile(travel_time, 0.75))
}

cache('station_weights')
```

#### Combine station ids
```{r}
station_route_ids <- unique_sequences %>% 
  select(route_id, stop_id, stop_name, prev_stop_id, prev_stop_name, direction_id) %>% distinct() %>% filter(! is.na(prev_stop_id)) %>%
  spread(key = route_id, value = 1) %>% 
  unite("route_ids", -stop_id, -prev_stop_id, -direction_id, -stop_name, -prev_stop_name) %>%
  mutate(route_ids = gsub("(_NA)|(NA_)","", route_ids))
```

#### Combine realtime with static edges
```{r}
# Static with realtime weights and route_ids
full_sequences <- station_route_ids %>%
  left_join(station_weights, by = c("stop_id" = "stop_mta_id", "prev_stop_id" =
                              "prev_stop_mta_id"))%>%
  select(route_ids, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name, 
         weight, sd, lower_quartile, median, upper_quartile)
  
# Transfers already has weights from the min_transfer_time field (now renamed weight) so we rbind the two sequence objects
full_sequences <- bind_rows(full_sequences, transfer_sequences)

```

#### Make an igraph
```{r igraph}

theme_set(theme_bw())

# Input a dataframe of the edges that exist between stops, with weight and direction as attributes
igraph_edges <- full_sequences %>% ungroup() %>%
  filter(!is.na(prev_stop_id), !is.na(weight)) %>%
  select(prev_stop_id, stop_id, weight, route_ids, direction_id)

mta_igraph <- graph.data.frame(igraph_edges, directed=TRUE)

```

#### Function to get formatted itineraries
```{r}
get_itinerary <- function(shortest_paths_df, num_ititeraries) {
  
  # new df for the formatted itineraries
  itinerary <- setNames(data.frame(matrix(ncol = 7, nrow = 0)),
                              c("itinerary_id", "station", "line", "direction", "leg", "event", "event_id"))
  
  # format each itinerary_id separately
  for (i in 1:num_ititeraries) {
    df <- shortest_paths_df %>%
      filter(itinerary_id == i)
    
    # current index
    k = 1
    
    # keep track of leg
    l = 1
    
    df <- df %>% mutate(leg = l, event = "travel")
    df$event[1] <- "start_trip"
    
    # compare current and next rows for whole itinerary
    while (k < nrow(df)){
      # next index
      j = k+1
      
      # identify transfers
      if (df$station[k] != df$station[j] & df$line[k] != df$line[j]) {
        
        # identify an 'implicit transfer (e.g. transfer 120->120 from 1 to 2/3)
        if (df$line[j] != "") {
          df <- df %>% add_row(itinerary_id = df$itinerary_id[j], station = df$station[j],
                line = df$line[k], direction = df$direction[k], leg = l, .after = k)
        }
        
        # identify an 'explicit' transfer (e.g. transfer R11->629 from N/R to 4)
        else {
          df$line[j] <- df$line[k] 
          df$leg[j] <- df$leg[k]
        }
        
        # make note of transfer events
        df$event[j] <- "start_transfer"
        df$event[j+1] <- "end_transfer"
        
        # start_transfer direction gets inherited from previous station
        df$direction[j] <- df$direction[k]
        
        # skip newly added row
        k = k+1
        
        # new leg starts with a transfer
        l = l+1

      }
      
      k = k+1
      df$leg[k] <- l
      
    }
    
    # very last row is redundant
    df <- df[1:nrow(df)-1,]
    df$event[nrow(df)] <- "end_trip" 
    df <- df %>% mutate(event_id = seq(1:nrow(df)))
  
    # put it all together
    itinerary <- rbind(itinerary, df) 
  }
  itinerary <- itinerary %>% left_join(stops, by = c("station" = "stop_id")) %>% select(itinerary_id, station, stop_name, line, direction, event_id, event, leg)

    return(itinerary)
}
```

#### format data for map
```{r}
format_map_data <- function(itin){
  routes <- routes_data
  routes$route_color <- replace_na(routes$route_color, "000000")
  routes <- routes %>% mutate(route_color = sprintf("#%s", route_color))
 
   itin_map_data <- itin %>% left_join(stops, by = c("station" = "stop_id", 
                                                    "stop_name" = "stop_name"))
  itin_map_data <- itin_map_data %>% 
    extract(line, c("route_id"),"([^_][^_]?)_?.*", remove = FALSE) %>% 
    left_join(routes)
  
  return(itin_map_data)
}
```

#### Multiple maps from multiple paths
```{r}
leaflet_multiple_maps <- function(itin_map_data, k_num){
  maps = list()
  for(j in 1:k_num){
    itin_map <- itin_map_data %>% filter(itinerary_id == j)
    map <- leaflet() %>%
      addTiles() %>%
      setView(-74.00, 40.71, zoom = 12) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addCircleMarkers(itin_map$stop_lon, itin_map$stop_lat, 
                       color = itin_map$route_color,
                       popup = itin_map$stop_name,
                       radius = 3) 
    
    for (i in 1:(nrow(itin_map)-1)) {
      map <- map %>%
        addPolylines(lat = c(itin_map[i,]$stop_lat, itin_map[i+1,]$stop_lat),
                     lng = c(itin_map[i,]$stop_lon, itin_map[i+1,]$stop_lon),
                     color = itin_map[i,]$route_color,
                     popup = itin_map[i,]$line,
                     dashArray = ifelse(itin_map[i,]$event == "start_transfer",
                                        c(8), c(0)), weight = 1)
    }
    #Add map to list
    maps[[j]] = map
  }
  return(maps)
}
```

#### Test itinerary function
```{r}
k_num <- 5

#Find k_num shortest paths leading going from stop R13 to stop D05 
k_shortest <- k_shortest_path(mta_igraph, "A27", "132", k=k_num)
itin <- get_itinerary(k_shortest, k_num)

map_data <- format_map_data(itin)
sync(leaflet_multiple_maps(map_data, k_num))
```