---
title: "Itinerary (hopefully final)"
author: "Sasha, Ayliana, Taxi"
date: "July 20, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Sasha, Ayliana, Brian, & Taxi*  

#### Load libraries
```{r load-libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(igraph)
library(lubridate)
library(leaflet)
library(mapview)
```

#### Load data
```{r load-data, warning=FALSE, message=FALSE}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
route <- read_csv('../../data/google_transit_subway_static/routes.txt')

# time/day filtering happens in '../../data/get_igraph.R'
# if necessary, change filters there and rerun script before running next line
load('../../data/mta_igraph.RData')

```


### *Functions*
#### Fixed k_shortest
```{r fixed_k_shortest}
#returns the shortest path from to 'to'
#when there is no path returns null instead of warning message
#also instead of returning vertex id, it returns vertex name
shortestPath <- function(graph, from, to){
  path <- tryCatch(get.shortest.paths(graph,from, to, mode = 'out', output = 'vpath'), warning = function(x) NULL)
  names(unlist(path$vpath))
}

#calculates the distance/length/sum of weights for a given path
#the path should be a list of vertex names. e.g c('a','b','c') 
distance <- function(graph, path) sum(E(graph, path=path)$weight)

#sorts the path
sortPath <- function(graph,paths) paths[paths %>% sapply(distance, graph=graph) %>% order]

#yen
k_shortest.yen <- function(graph, from, to, k){
  A <- list(shortestPath(graph,from,to))
  B <- list()
  for (k_i in 2:k){
    for(i in 1:(length(A[[k_i-1]])-1)){
      spurNode <- A[[k_i-1]][i]
      rootPath <- A[[k_i-1]][1:i]
      edgesToDelete <- list()
      for (p in A){
        if (all(p[1:i] == rootPath)) {
          edge <- paste(p[i], p[i+1], sep = '|')
          edgesToDelete <- append(edgesToDelete, edge)
        }
      }
      edgesToDelete <- unique(edgesToDelete)
      
      t_g <- graph
      for (edge in edgesToDelete){
        t_g <- delete.edges(t_g, edge)
      }
      spurPath <- shortestPath(t_g,spurNode,to)
      if (!is.null(spurPath)){
        total_path <- list(c(rootPath[-i], spurPath))
        if (!total_path %in% B) B[length(B)+1] <- total_path
        #print('fouuuuuund')
      }
      #print(paste('I:', i))
      #print(paste('A:',paste(A[[k_i-1]], collapse = ' ')))
      #print(paste('Edges TO Remove: ', paste(edgesToDelete, collapse = ' ')))
      #print(paste('spurNode', spurNode))
      #print(paste('spurPath:', paste(spurPath, collapse = ' ')))
      #print(paste('RootPath:', paste(rootPath, collapse = ' ')))
      #print(paste('TotalPath: ', paste(total_path, collapse = ' ')))
      #print('========================')
    }
    if (length(B) == 0) break
    B <- sortPath(graph, B)
    A[k_i] <- B[1]
    B <- B[-1]
   # print('***************************************')
  }
  A
}

#extracts route_ids, and direction_id from path
extract_data <- function(graph,i,path){
  edges <- E(graph, path=path)
  
  direction <- edges$direction_id
  direction <- c(direction,direction[length(direction)])
  
  line <- c(edges$route_ids,'end')
  
  tibble(itinerary_id = i, station = path, line, direction)
}
#combines paths into a tibble
paths_to_tibble <- function(graph, paths) {
  paths.asTibble <- 1:length(paths) %>%
  lapply(function(i) extract_data(graph,i, paths[[i]])) %>%
  reduce(rbind) 
}

#wrapper around yen, processes yen's result so the itinery functon can interact with it
k_shortest_path <- function(graph, from, to, k) k_shortest.yen(graph, from, to, k) %>% paths_to_tibble(graph=graph)
```


#### Greedy function
```{r}
greedy <- function(shortest_paths_df, num_itineraries){
  all_lines <- vector(mode = "character")
  
  for(i in 1:num_itineraries){
    
    df <- shortest_paths_df %>% 
      filter(itinerary_id == i)
    
    lines <- vector(mode = "character")
    
    # keep track of how many prev_line_ids to append
    r = 0
    
    prev_line_ids <-  str_split(df$line[i], "_")[[1]]
    
    for (j in 1:nrow(df)) {
      
      cur_line_ids <- str_split(df$line[j], "_")[[1]]
      intersect_lines <- intersect(cur_line_ids, prev_line_ids)
      
      if (length(intersect_lines) == 0){
        
        if("T" %in% cur_line_ids){
          
          shortest_paths_df[j,]$direction <- shortest_paths_df[j-1,]$direction
        }
        
        else{
          # otherwise set to current line_ids
          intersect_lines <- cur_line_ids
        }
      
        # fill in all of the common lines up until the transfer
        lines <- append(lines, rep(paste(prev_line_ids, collapse = "_"), r))
        r = 0
      }
 
      r = r+1
      prev_line_ids <- intersect_lines
    
      }
   
    all_lines <- append(all_lines, lines)
    all_lines <- append(all_lines, "END")
    
  } # endfor
  shortest_paths_df$line <- all_lines
  
  return(shortest_paths_df)

}

```


#### Function to get formatted itineraries
```{r}

get_itinerary <- function(shortest_paths_df, num_ititeraries) {
  
  # new df for the formatted itineraries
  itinerary <- setNames(data.frame(matrix(ncol = 7, nrow = 0)),
                              c("itinerary_id", "station", "line", "direction", "leg", "event", "event_id"))
  
  # get correct lines with greedy function
  shortest_paths_df <- greedy(shortest_paths_df, num_ititeraries)
  
  # format each itinerary_id separately
  for (i in 1:num_ititeraries) {
    df <- shortest_paths_df %>%
      filter(itinerary_id == i)
    
    # current index
    k = 1
    
    # keep track of leg
    l = 1
    
    df <- df %>% mutate(leg = l, event = "travel")
    df$event[1] <- "start_trip"
    
    # compare current and next rows for whole itinerary
    while (k < nrow(df)){
      # next index
      j = k+1
      
      # identify transfers
      if (df$station[k] != df$station[j] & df$line[k] != df$line[j]) {
        
        # identify an 'implicit transfer (e.g. transfer 120->120 from 1 to 2/3)
        if (df$line[j] != "") {
          df <- df %>% add_row(itinerary_id = df$itinerary_id[j], station = df$station[j],
                line = df$line[k], direction = df$direction[k], leg = l, .after = k)
        }
        
        # identify an 'explicit' transfer (e.g. transfer R11->629 from N/R to 4)
        else {
          df$line[j] <- df$line[k] 
          df$leg[j] <- df$leg[k]
        }
        
        # make note of transfer events
        df$event[j] <- "start_transfer"
        df$event[j+1] <- "end_transfer"
        
        # start_transfer direction gets inherited from previous station
        df$direction[j] <- df$direction[k]
        
        # skip newly added row
        k = k+1
        
        # new leg starts with a transfer
        l = l+1

      }
      
      k = k+1
      df$leg[k] <- l
      
    }
    
    # very last row is redundant
    df <- df[1:nrow(df)-1,]
    df$event[nrow(df)] <- "end_trip" 
    df <- df %>% mutate(event_id = seq(1:nrow(df)))
  
    # put it all together
    itinerary <- rbind(itinerary, df)
  }
    itinerary <- itinerary %>%
      left_join((stops %>% select(stop_id, stop_name)), by=c("station" = "stop_id"))
    
    return(itinerary)
}
```


#### Get coordinates for the map
```{r get-map-coords}

# take care of NA color values
routes <- route
routes$route_color <- replace_na(routes$route_color, "000000") 
routes <- routes %>%
  mutate(route_color = sprintf("#%s", route_color))

# retrieve coordinates and stop names for itineraries
get_map_data <- function(itinerary) {
  
  map_data <- itinerary %>%
  left_join(stops, by = c("station" = "stop_id", "stop_name" = "stop_name"))  %>%
  group_by(itinerary_id) %>%
  mutate(prev_station = lag(station), prev_stop_name = lag(stop_name),
         prev_stop_lat = lag(stop_lat), prev_stop_lon = lag(stop_lon)) %>%
  extract(line, c("route_id"), regex="([^_][^_]?)_?.*", remove=FALSE) %>%
  ungroup() %>%
  left_join(routes) %>%
  select(itinerary_id, station, line, direction, leg, event, event_id,
         stop_name, stop_lat, stop_lon, prev_station, prev_stop_name, prev_stop_lat,
         prev_stop_lon, route_color)
  
  return(map_data)
  
}
```


#### Function to map itineraries (as different layers)
```{r layered-map-function}
# map a map with leaflet
map_an_itinerary <- function(itinerary) {
  
  black <- "#000000"
  
  map_df <- get_map_data(itinerary) %>%
    mutate(line = str_replace_all(line, "_", " "))
  
  map_df %>% View
  
  num_layers <- max(map_df$itinerary_id)

  map <- leaflet() %>%
    addTiles() %>%
    setView(median(map_df$stop_lon), median(map_df$stop_lat), zoom = 12) %>%
    addProviderTiles("CartoDB.Positron")

for (i in 1:num_layers) {
  df <- map_df %>% filter(itinerary_id==i)

  map <- map %>%
    addCircleMarkers(df$stop_lon, df$stop_lat, color = df$route_color,
                     popup = paste("<b>", df$stop_name, "</b>", "<br/>", df$line),
                     radius = 3,
                     group= as.character(i))

  for (j in 1:nrow(df)) {
    map <- map %>%
      addPolylines(lat = c(df[j,]$stop_lat, df[j,]$prev_stop_lat),
                   lng = c(df[j,]$stop_lon, df[j,]$prev_stop_lon),
                   color = ifelse(df[j,]$event == "end_transfer", black, df[j,]$route_color),
                   popup = ifelse(df[j,]$event == "end_transfer",
                                  paste(df[j-1,]$line, "to", df[j,]$line), df[j,]$line),
                   dashArray = ifelse(df[j,]$event == "end_transfer", c(8), c(0)),
                   weight = 2,
                   group = as.character(i))
  }

}

  map <- map %>%
    addLayersControl(overlayGroups = as.character(seq(1:num_layers)),
                     options = layersControlOptions(collapsed = FALSE))
  return(map)
}

```


#### Function to map itineraries (as synced tiles)
```{r tile-maps, message=FALSE}
tile_maps <- function(itinerary) {
  black <- "#000000"
  maps = list()
  map_df <- get_map_data(itinerary) %>%
      mutate(line = str_replace_all(line, "_", " "))
  
  for(j in 1:k_num){
    df <- map_df %>% filter(itinerary_id == j)
    
    map <- leaflet() %>%
      addTiles() %>%
      setView(median(map_df$stop_lon), median(map_df$stop_lat), zoom = 11) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addCircleMarkers(df$stop_lon, df$stop_lat, 
                       color = df$route_color,
                       popup = paste("<b>", df$stop_name, "</b>", "<br/>", df$line),
                       radius = 3) 
    
    for (i in 1:nrow(df)){
      map <- map %>%
        addPolylines(lat = c(df[i,]$stop_lat, df[i,]$prev_stop_lat),
                   lng = c(df[i,]$stop_lon, df[i,]$prev_stop_lon),
                   color = ifelse(df[i,]$event == "end_transfer", black, df[i,]$route_color),
                   popup = ifelse(df[i,]$event == "end_transfer",
                                  paste(df[i-1,]$line, "to", df[i,]$line), df[i,]$line),
                   dashArray = ifelse(df[i,]$event == "end_transfer", c(8), c(0)),
                   weight = 2,
                   group = as.character(i))
    }
    
    #Add map to list
    maps[[j]] = map
  }
  
  sync(maps)
  
}
```


### *Tests*
#### Run k shortest paths function
```{r}
k_num <- 5

# find k_num shortest paths leading going from stop R13 to stop D05 
k_shortest <- k_shortest_path(mta_igraph, from = "R13", to = "D05", k=k_num)

as.data.frame(k_shortest)
```


#### Get itinerary!
```{r test-get-itin}
itin <- get_itinerary(k_shortest, k_num)

as.data.frame(itin)
```


#### **Map it!**
#### View as a map with layers
```{r test-get-map, message=FALSE}
# get_map_data(itin) %>% View
map_an_itinerary(itin)

```


#### View as synced tiled maps
```{r test-tile-map, message=FALSE}
tile_maps(itin)
```
