---
title: "Commutes & Census"
author: "Sasha, Ayliana, Taxi"
date: "July 30, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(RColorBrewer)
library(sp)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(mapview)
library(ggplot2)
library(scales)
library(leaflet)
library(tidyverse)
library(ggrepel)
```

```{r load data}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
stops <- stops %>% dplyr::select(stop_id, stop_lat, stop_lon, stop_name)
routes_data <- read_csv('../../data/google_transit_subway_static/routes.txt')
source('../../src/path_finding.R')
load('../../data/igraph_edges.rdata')
load('../../data/income_merged.Rdata')
load('../../data/wait_times_rush_hour.rdata')

r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)
```

#### distance function
```{r}
distance <- function(graph, path) sum(E(graph, path=path)$weight)
```

#### geodistance (in km)
```{r}
Distance <- function(LatA, LonA, LatB, LonB){
  scaling_factor <- 110.25
  x <- LatA - LatB
  y <- (LonA - LonB) * cos(40.75 * (pi)/180)

  return (scaling_factor*sqrt(x * x + y * y))

}
```



```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


#### get graph
```{r}
igraph_edges <- mutate(igraph_edges, weight = `50%`)
igraph_edges <- igraph_edges %>% dplyr::select(stop_id, nxt_stop_id, weight, route_id, `90%`)
igraph_edges$weight <- as.numeric(igraph_edges$weight)
igraph_edges$`90%` <- as.numeric(igraph_edges$`90%`)
mta_igraph <- graph_from_data_frame(igraph_edges)
```


```{r message=FALSE, warning=FALSE}
to <- "128"


vertices <- names(V(mta_igraph))
vertices <- vertices[vertices != to]
station_list <- list()
time_50 <- list()
time_90 <- list()
for(i in 1:length(vertices)){
  path <- get_itinerary(mta_igraph, vertices[i], to, 1, stops = stops,
                                 attributeNames = '90%')
  station_list[i] <- vertices[i]
  path$weight <- as.numeric(path$weight)
  time_50[i] <- sum(path$weight, na.rm = TRUE)
  path$`90%` <- as.numeric(path$`90%`)
  time_90[i] <- sum(path$`90%`, na.rm = TRUE)
}

map_data <- data.frame(station = unlist(station_list))
map_data$time_50 <- unlist(time_50)
map_data$time_90 <- unlist(time_90)
map_data <- map_data %>% left_join(stops, by = c("station" = "stop_id")) %>%
  filter(! is.na(stop_lat))
map_data <- map_data %>% left_join(wait_times_filter, by = c("station" = "stop_mta_id")) %>% mutate(time_50 = time_50 + wait_time_median, time_90 = time_90 + wait_time_90)
```


#### Create a grid of points
```{r make-grid}
map_data_lats <- seq(min(map_data$stop_lat), max(map_data$stop_lat), by=0.003)
map_data_lons <- seq(min(map_data$stop_lon), max(map_data$stop_lon), by=0.003)
grid <- expand.grid(map_data_lats, map_data_lons) %>%
 dplyr::select(lat = Var1, lon = Var2)
```


#### Get subway travel time, walking time, total time
##### Filter for points reachable within 20 mins of walking
```{r get-times}
time_50 = vector()
time_90 = vector()
station = vector()
walking_time = vector()
total_time = vector()
for(i in 1:nrow(grid)){
 result<- map_data %>% mutate(walking_time = 720*Distance(grid$lat[i], grid$lon[i], stop_lat, stop_lon), total_time = (walking_time + time_50))
 idx <- which.min(result$total_time)
 
 
 time_50[i] <- result[idx, "time_50"]
 time_90[i] <- result[idx, "time_90"]
 station[i] <- result[idx, "stop_name"]
 walking_time[i] <- result[idx, "walking_time"]
 total_time[i] <- result[idx, "total_time"]
}
grid$time_50P <- time_50
grid$time_90P<- time_90
grid$nearest_stop <- station
grid$walking_time <- walking_time
grid$total_time <- total_time
#minute diff
grid <- grid %>% mutate(time_diff = ((time_90P-time_50P) / 60), walking_mins = (walking_time / 60))

#pct diff
# grid$diff <- grid$time_90P-grid$time_50P
# grid <- grid %>% mutate(time_diff = round(diff/total_time*100,2), walking_mins = (walking_time / 60))
grid$log_mins <- log(grid$time_diff)

```



### CENSUS
#### join with census
```{r spatial-join}

filtered_grid <- grid %>%
  filter(!is.na(time_diff), time_50 != 0) %>% #filter out transfers
  filter(walking_mins <= 20) %>%
  mutate(min_interval = cut(time_diff, seq(0, max(time_diff), by = 5), include.lowest = TRUE),
         bin = cut(time_diff, seq(0, max(time_diff), by = 5), include.lowest = TRUE, labels = FALSE))

filtered_grid_spdf <- filtered_grid
coordinates(filtered_grid_spdf) <- ~lon + lat
proj4string(filtered_grid_spdf) <- proj4string(income_merged)
matches <- over(filtered_grid_spdf, income_merged)
filtered_grid <- cbind(filtered_grid, matches)
filtered_grid <- filtered_grid %>%
  filter(!is.na(GEOID), !is.na(median_income))
coords <- filtered_grid[c("lon", "lat")]

filtered_grid_spdf <- filtered_grid
coordinates(filtered_grid_spdf) <- ~lon + lat
proj4string(filtered_grid_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(filtered_grid_spdf, nyc_neighborhoods)
filtered_grid <- cbind(filtered_grid, matches)
filtered_grid <- filtered_grid %>%
  filter(!is.na(neighborhood))
coords <- filtered_grid[c("lon", "lat")]

```


```{r}
ggplot(filtered_grid, aes(x = median_income, y = time_diff)) +
  geom_point(aes(color = borough)) +
  xlab('Median Household Income') +
  scale_x_continuous(label = comma) +
  ylab('Difference in Commute Time to 23rd St') + 
  scale_color_discrete("Borough")

ggplot(filtered_grid, aes(x = median_income, y = time_90P, color = borough)) +
  geom_point(alpha = 0.2) +
  xlab('Median Household Income') +
  scale_x_continuous(label = comma) +
  ylab('time_90P to 34th St') + 
  scale_color_discrete("Borough")+
  geom_smooth(method = "lm") +
  facet_wrap(~borough)

plot_90 <- ggplot(filtered_grid, aes(x = median_income, y = time_90P, color = borough)) +
  geom_point(alpha = 0.2) +
  xlab('Median Household Income') +
  scale_x_continuous(label = comma) +
  ylab('90th Perc. Trip Time to 34th St') + 
  scale_color_discrete("Borough")+
  geom_smooth(method = "lm") +
  facet_wrap(~borough)
```


#### for the bronx, and neighborhood with median income around 50k, (try to) make a frontier plot
```{r}
filtered_grid %>% 
  mutate(median_income = pmin(round(median_income/25000)*25000, 100000),
         time_variance = time_90P - time_50P,
         time_50P_mins = time_50P/60,
         time_variance_mins = time_variance/60) %>% 
  filter(borough == "Bronx", median_income == 50000) %>%
  ggplot(aes(x = time_50P/60, y = time_90P/60 - time_50P/60, color=nearest_stop, label = nearest_stop)) +
  geom_point(alpha = 0.2) +
  geom_text(position = position_jitter(3)) +
  # geom_text_repel() +
  xlab('Median Trip Time to 34th') +
  scale_x_continuous(label = comma) +
  ylab('90th - 50th Perc. Trip Time to 34th St') +
  # facet_wrap(~neighborhood)
  scale_color_discrete(guide=FALSE)
  # # geom_smooth(method = "lm") +
  # facet_grid(median_income~borough, label=comma)
```




### start here to replicate slide deck figures
#### plot income vs commute time for brooklyn, manhattan, queens
```{r}

filtered_grid_all <- filtered_grid %>%
  filter(borough != "Staten Island", borough != "Bronx")

plot_50 <- ggplot(filtered_grid_all, aes(x = median_income, y = time_50P/60 + walking_mins, color = borough)) +
  geom_point(alpha = 0.2) +
  xlab('Median Household Income') +
  scale_x_continuous(label = comma) +
  ylab('Median Trip Time to 34 St') +
  scale_y_continuous(label = comma) +
  scale_color_discrete("Borough")+
  geom_smooth(method = "lm") +
  facet_wrap(~borough) +
  theme_grey(base_size = 20) +
  theme(legend.position = "bottom")

plot_50

ggsave("income_v_50P_time_to_34.png", plot = plot_50, height = 6, width = 18)
```


#### plot income vs commute time for bronx
```{r}
filtered_grid_bx <- filtered_grid %>%
  filter(borough == "Bronx")

plot_50 <- ggplot(filtered_grid_bx, aes(x = median_income, y = time_50P/60 + walking_mins, color = borough)) +
  geom_point(alpha = 0.2, color = "purple") +
  xlab('Median Household Income') +
  scale_x_continuous(label = comma) +
  ylab('Median Trip Time to 34 St') +
  scale_y_continuous(label = comma) +
  scale_color_discrete("Borough")+
  geom_smooth(method = "lm", color = "purple") +
  facet_wrap(~borough) +
  theme_grey(base_size = 20) +
  theme(legend.position = "bottom")

plot_50

ggsave("income_bronx.png", plot = plot_50, height = 7, width = 7)
```



#### map higher income neighborhoods in the bronx
```{r}
manhattan_all <- filtered_grid %>% filter(borough == "Manhattan")
bronx_all <- filtered_grid %>% filter(borough == "Bronx")
bronx_75 <- filtered_grid %>% filter(borough == "Bronx", median_income >= 75000) 
bronx_income <- leaflet(nyc_neighborhoods) %>%
      addTiles() %>%
      addProviderTiles("CartoDB.Positron") %>%
      setView(-73.87, 40.81, zoom = 12) %>%
  addCircleMarkers(data = manhattan_all, lat = ~lat, lng = ~lon, 
                   radius = 6, stroke = FALSE, color = "#b0d9bf", fillOpacity = 0.4) %>%
  addCircleMarkers(data = bronx_all, lat = ~lat, lng = ~lon, 
                   radius = 6, stroke = FALSE, color = "#d4c1de", fillOpacity = 0.4) %>%
  addCircleMarkers(data = bronx_75, lat = ~lat, lng = ~lon, 
                   radius = 6, stroke = FALSE, color = "purple", fillOpacity = 0.7)

bronx_income

mapshot(bronx_income, url = "bronx_income_75000.html", 
                                       file ="bronx_income_75000.png", vheight = 1000, vwidth = 1000)

```
