---
title: "2nd Ave"
author: "Sasha, Ayliana, Taxi"
date: "July 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries
```{r libs, warning=FALSE, message=FALSE}
library(igraph)
library(leaflet)
library(RColorBrewer)
library(geosphere)
library(sp)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(ggmap)
library(ggplot2)
library(gstat)
library(raster)
library(mapview)
library(tidyverse)
library(profvis)
```

#### Load data
```{r load-data, message= FALSE}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
stops <- stops %>% select(stop_id, stop_lat, stop_lon, stop_name)

routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
routes$route_color <- replace_na(routes$route_color, "000000") 
routes <- routes %>%
  mutate(route_color = sprintf("#%s", route_color))

source('../../src/path_finding.R')
load('../../data/igraph_edges.rdata')

r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)
```


#### distance function for summing graph weights
```{r dist1}
distance <- function(graph, path) sum(E(graph, path=path)$weight)
```



#### distance function for walking to stations (in km)
```{r dist2}
Distance <- function(LatA, LonA, LatB, LonB){
  scaling_factor <- 110.25
  x <- LatA - LatB
  y <- (LonA - LonB) * cos(40.75 * (pi)/180)

  return (scaling_factor*sqrt(x * x + y * y))

}
```



#### heatmap stuff
```{r make-grid}

# get time from each subway station to "dest"
get_heatmap_static_info <- function(dest, igraph) {
  map_data <- get.all.shortest.paths(igraph, dest,  mode = "out")

  map_data <- map_data$res %>% 
    lapply(function(x)data.frame(stop_id = names(x)[length(x)],
                                 distance <- distance(igraph, x), 
                                 path = paste0(names(x), collapse = '>'))) %>%
    reduce(rbind) 

  map_data <- map_data %>% left_join(stops) 
  names(map_data) <- c("stop_id", "distance", "path", "stop_lat", "stop_lon", "stop_name")
  

  map_data <- map_data %>%
    mutate(subway_mins = round(distance/60, 2))
  
  map_data$path <- as.character(map_data$path)
  
  return(map_data)

}
```


#### Functions for making commute heatmaps given 1 destination
```{r regular}
# get time from each point on grid to "dest" 
get_heatmap_grid_info <- function(map_data, grid) {
  total_time = vector()
  walking = vector()
  station = vector()
  sequence = vector()
  
  for(i in 1:nrow(grid)){
    result<- map_data %>% mutate(walking_time = 720*Distance(grid$lat[i], grid$lon[i], stop_lat, stop_lon), 
                                 total_time = (walking_time + distance))
    idx <- which.min(result$total_time)
    
    total_time[i] <- result[idx, "total_time"]
    walking[i] <- result[idx, "walking_time"]
    station[i] <- result[idx, "stop_name"]
    sequence[i] <- result[idx, "path"]
    # print(walking[i])
  }
  grid$time <- total_time
  grid$walking <- walking
  grid$nearest_stop <- station
  grid$path <- sequence
  grid <- grid %>% mutate(total_time_mins = round(time/60, 2), walking_mins = round(walking/60, 2), subway_mins = total_time_mins - walking_mins)
  
  return(grid) 
}

# changes depending on variable being measured  
get_palette <- function(grid) {
  # filter for trips with 20 mins or less of walking
  filtered_grid <- grid %>%
    filter(walking_mins <= 20) %>%
    mutate(min_interval = cut(total_time_mins, seq(0, max(total_time_mins), by = 5), include.lowest = TRUE),
           bin = cut(total_time_mins, seq(0, max(total_time_mins), by = 5), include.lowest = TRUE, labels = FALSE))
  
  clrs <- brewer.pal(11, "RdYlGn")
  
  grid_pal <- colorNumeric(
    palette = clrs,
    domain = sqrt(filtered_grid$total_time_mins), #doesn't have to be sqrt()
    reverse = TRUE,
    na.color = "transparent")
  
  return(grid_pal)
}
```



#### Calculate differences in total, subway, and walking times with/without the Q
```{r}
get_diffs <- function(grid1, grid2) {
  all_times <- left_join(grid1, grid2, by=c("lat", "lon"))
  time_diffs <- all_times %>%
    filter(path.x != path.y) %>%
    mutate(total_diff = total_time_mins.y - total_time_mins.x,
           subway_diff = total_time_mins.y - total_time_mins.x,
           walking_diff = walking_mins.y - walking_mins.x)
  
  return(time_diffs)
}

```




#### Map the difference in total time
```{r}

get_diff_heatmaps <- function(time_diffs, diffs = c("total_diff", "subway_diff", "walking_diff")) {

  maps = list()
  for (i in 1:length(diffs)) {
    filtered_grid <- time_diffs %>%
      filter(walking_mins.x <= 30 | walking_mins.y <= 30)
    
    filtered_grid_spdf <- filtered_grid
    coordinates(filtered_grid_spdf) <- ~lon + lat
    proj4string(filtered_grid_spdf) <- proj4string(nyc_neighborhoods)
    matches <- over(filtered_grid_spdf, nyc_neighborhoods)
    filtered_grid <- cbind(filtered_grid, matches)
    filtered_grid <- filtered_grid %>%
      filter(!is.na(neighborhood))

    coords <- filtered_grid[c("lon", "lat")]

    gridded(filtered_grid_spdf) <- TRUE

    idw <- idw(formula=filtered_grid_spdf[[diffs[i]]] ~ 1, locations=filtered_grid_spdf, newdata = filtered_grid_spdf)

    residual_grid <- raster(idw, "var1.pred")
    # contourr <- rasterToContour(residual_grid)

    rr <- mask(residual_grid, nyc_neighborhoods)
    crs(rr) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

    clrs <- brewer.pal(11, "RdYlGn")

    grid_pal <- colorNumeric(
      palette = clrs,
      domain = filtered_grid[[diffs[i]]],
      reverse = TRUE,
      na.color = "transparent")

    m <- leaflet(nyc_neighborhoods) %>%
      addTiles() %>%
      addProviderTiles("CartoDB.Positron") %>%
      setView(-73.87, 40.75, zoom = 11) %>%
      addLegend("bottomright", pal = grid_pal, values = filtered_grid[[diffs[i]]],
                title = paste("Change in Commute Time", "<br/>", diffs[i]),
                labFormat = labelFormat(suffix = " mins"),
                opacity = 0.5) %>%
      addRasterImage(rr, color=grid_pal, opacity = 0.5)

    maps[[i]] <- m
  }

  sync(maps)

}
```




#### settings
```{r}
# final destination station
dest <- "A31" # 14th St ******
# dest <- "103"
dest <- "631" #grand central
dest <- "128" #penn

```


### GET GRAPHS
#### with L
```{r}
igraph_edges <- mutate(igraph_edges, 'weight'=`90%`)
graph <- graph.data.frame(igraph_edges)
```

#### without L
```{r}
second_ave <- c("Q03", "Q04", "Q05")

no_q_edges <- igraph_edges %>%
  filter((!stop_id %in% second_ave) & (!nxt_stop_id %in% second_ave))
q_graph <-graph.data.frame(no_q_edges)

```



#### **need this for heatmap**
#### make a grid of points over nyc
```{r}
grid_lats <- seq(min(stops$stop_lat), max(stops$stop_lat), by=0.002)
grid_lons <- seq(min(stops$stop_lon), max(stops$stop_lon), by=0.002)
grid <- expand.grid(grid_lats, grid_lons) %>%
    select(lat = Var1, lon = Var2)
```



```{r}
graph_static <- get_heatmap_static_info(dest, graph)
graph_grid <- get_heatmap_grid_info(graph_static, grid)

# 1 min for 0.002 grid
# 15 s for 0.004 grid


q_graph_static <- get_heatmap_static_info(dest, q_graph)
q_graph_grid <- get_heatmap_grid_info(q_graph_static, grid)

time_diffs <- get_diffs(graph_grid, q_graph_grid)

get_diff_heatmaps(time_diffs)


# 16 s for 0.006 grid
# 22 s for 0.005 grid
# everything: 32 s for 0.004 grid
# 2 mins for 0.002 grid

```

