---
title: "Maps"
author: "Sasha, Ayliana, Taxi"
date: "July 27, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Load libraries
```{r warning=FALSE, message=FALSE}
library(igraph)
library(leaflet)
library(RColorBrewer)
library(geosphere)
library(sp)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(ggmap)
library(KernSmooth)
library(ggplot2)
library(gstat)
library(raster)
library(mapview)
library(tidyverse)
```

#### Load data
```{r load-data, message= FALSE}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt') %>% 
  select(stop_id, stop_lat, stop_lon, stop_name)

routes <- read_csv('../../data/google_transit_subway_static/routes.txt') %>% 
  mutate(route_color = replace_na(route_color, '000000'), route_color = sprintf("#%s", route_color))

load('../../data/igraph_edges.rdata')

accessible_stations <- read_csv('../../data/accessible_stations.csv') [[1]]

r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)

stations <- read_csv('http://web.mta.info/developers/data/nyct/subway/Stations.csv')
source('../../src/path_finding.R')
source('disability.r')
load('./heatmap_summary')
```


### GET GRAPHS
```{r}
graph_50 <- mutate(igraph_edges, weight = `50%`) %>% graph.data.frame()

graph_90 <- mutate(igraph_edges, weight =`90%`) %>% graph.data.frame()

graph_mean <- mutate(igraph_edges, weight = mean) %>% graph.data.frame()
```


#### **need this for heatmap**
#### make a grid of points over nyc
```{r}
grid <- create_grid(stops$stop_lat, stops$stop_lon,nyc_neighborhoods)
```


#### make station2station and point2station distance data for accessible stations only
```{r regular}
dest = 'A31'
manual_wheelchair_speed = 15
station2station_90_accessible <- distance_station2station(graph_90,stops, dest, src = accessible_stations)
station2station_50_accessible <- distance_station2station(graph_50,stops, dest, src = accessible_stations)
station2station_mean_accessible <- distance_station2station(graph_mean,stops, dest, src = accessible_stations)

point2station_90_accessible <- distance_point2station(station2station_90_accessible, grid, manual_wheelchair_speed)
point2station_50_accessible <- distance_point2station(station2station_50_accessible, grid, manual_wheelchair_speed)
point2station_mean_accessible <- distance_point2station(station2station_mean_accessible, grid, manual_wheelchair_speed)
```

#### make station2station and point2station distance data for all stations
```{r regular}
station2station_90_all <- distance_station2station(graph_90,stops, dest)
station2station_50_all <- distance_station2station(graph_50,stops, dest)
station2station_mean_all <- distance_station2station(graph_mean,stops, dest)

point2station_90_all <- distance_point2station(station2station_90_all, grid)
point2station_50_all <- distance_point2station(station2station_50_all, grid)
point2station_mean_all <- distance_point2station(station2station_mean_all, grid)
```

#### make station2station and point2station distance data for all stations except for l train stations
```{r regular}
all_stops <- stations %>% pull('GTFS Stop ID')
graph_stops <- V(graph_90) %>% names()
diff <- graph_stops[!graph_stops %in% all_stops]
stops_without_l <- stations %>% filter(`Daytime Routes` != "L") %>% pull('GTFS Stop ID')
stops_without_l <- c(diff, stops_without_l[stops_without_l %in% graph_stops])

station2station_90_l <- distance_station2station(graph_90,stops, dest, stops_without_l)
station2station_50_l <- distance_station2station(graph_50,stops, dest, stops_without_l)
station2station_mean_l <- distance_station2station(graph_mean,stops, dest, stops_without_l)

point2station_90_l <- distance_point2station(station2station_90_l, grid)
point2station_50_l <- distance_point2station(station2station_50_l, grid)
point2station_mean_l <- distance_point2station(station2station_mean_l, grid)
```

#### maps with mean
```{r regular}

map_mean_accessible <- alternative_map(point2station_mean_accessible, dest, stops)
map_mean_all <- alternative_map(point2station_mean_all, dest,stops, paletteSource = point2station_mean_accessible)
map_mean_l <- alternative_map(point2station_mean_l, dest, stops, paletteSource = point2station_mean_accessible)

sync(map_mean_accessible, map_mean_all, map_mean_l)
point2station_mean_accessible_filtered <- point2station_mean_accessible %>% filter(total_time_mins < 120)
point2station_mean_all_filtered <- point2station_mean_all %>% filter(total_time_mins < 120)
point2station_mean_l_filtered <- point2station_mean_l %>% filter(total_time_mins < 120)

map_mean_accessible_filtered <- point2station_mean_accessible_filtered %>% 
  alternative_map(dest, stops)
map_mean_all_filtered <- point2station_mean_all_filtered %>% 
  alternative_map(dest, stops, paletteSource = point2station_mean_accessible_filtered)
map_mean_l_filtered <- point2station_mean_l_filtered %>%
  alternative_map(dest, stops, paletteSource = point2station_mean_accessible_filtered)

sync(map_mean_accessible_filtered, map_mean_all_filtered)
mapview::mapshot(map_mean_accessible_filtered, file = 'heatmap_accessible_filtered.png')
mapview::mapshot(map_mean_all_filtered, file = 'heatmap_regular_filtered.png')
#loop through all point and find the worst ones
```

#### diff
```{r}
diff_map(point2station_mean_accessible, point2station_mean_all,dest = dest, stops = stops, max = 40)
```

#### check stops with disability
```{r}
accessible_summary <- tibble()
for (dest in accessible_stations){
  temp <- distance_station2station(graph_mean, stops, dest, accessible_stations) %>% distance_point2station(grid,speed = 15)%>% 
    mutate(dest = dest) 
  accessible_summary<-bind_rows(accessible_summary, temp) 
}

all_summary <- tibble()
i <-1
for (dest in accessible_stations){
  print(i)
  i<-i+1
  temp <- distance_station2station(graph_mean, stops, dest) %>% distance_point2station(grid)%>% 
    mutate(dest = dest) 
  all_summary<-bind_rows(all_summary, temp) 
}
save(accessible_summary, all_summary, file = 'heatmap_summary')
```

```{r}

diff_summary <- left_join(accessible_summary, all_summary, by= c('lat','lon','dest')) %>% mutate(diff = total_time_mins.y -total_time_mins.x) %>% select(lon, lat, dest, diff) %>% group_by(dest) %>% summarize(avg = mean(diff), worst = max(diff), best = min(diff)) %>% arrange(avg)

best <- last(diff_summary$dest)
worst <- first(diff_summary$dest)
median <- diff_summary[nrow(diff_summary)/2, 'dest'][[1]]
q1 <- diff_summary[floor(nrow(diff_summary)*.25),'dest'][[1]]
q3 <- diff_summary[floor(nrow(diff_summary)*.75), 'dest'][[1]]



worst_point2station_accessible <- get_point2station(graph_90, stops, worst, src = accessible_stations, grid, speed = 15)
worst_point2station_all <- get_point2station(graph_90, stops, worst, grid = grid)

worst_point2station_accessible_filtered <- worst_point2station_accessible %>% filter(total_time_mins < 120)
worst_point2station_all_filtered <- worst_point2station_all %>% filter(total_time_mins <120) 

worst_map_accessible <- alternative_map(worst_point2station_accessible, worst, stops)
worst_map_all <- alternative_map(worst_point2station_all, worst, stops, paletteSource =  worst_point2station_accessible)
worst_diff_map <- diff_map(worst_point2station_accessible, worst_point2station_all, worst, stops)

worst_map_accessible_filtered <- alternative_map(worst_point2station_accessible_filtered, worst, stops)
worst_map_all_filtered <- alternative_map(worst_point2station_all_filtered, worst, stops, paletteSource =  worst_point2station_accessible_filtered)


q3_point2station_accessible <- get_point2station(graph_90, stops, q3, src = accessible_stations, grid, speed = 15)
q3_point2station_all <- get_point2station(graph_90, stops, q3, grid = grid)

q3_point2station_accessible_filtered <- q3_point2station_accessible %>% filter(total_time_mins < 120)
q3_point2station_all_filtered <- q3_point2station_all %>% filter(total_time_mins <120) 

q3_map_accessible <- alternative_map(q3_point2station_accessible, q3, stops)
q3_map_all <- alternative_map(q3_point2station_all, q3, stops, paletteSource =  q3_point2station_accessible)
q3_diff_map <- diff_map(q3_point2station_accessible, q3_point2station_all, q3, stops)

q3_map_accessible_filtered <- alternative_map(q3_point2station_accessible_filtered, q3, stops)
q3_map_all_filtered <- alternative_map(q3_point2station_all_filtered, q3, stops, paletteSource =  q3_point2station_accessible_filtered)
```



```{r}
get_point2station <- function(graph, stops, dest, src = V(graph), grid, speed = 12)
{
  distance_station2station(graph,stops, dest, src) %>%distance_point2station(grid, speed = speed)
}
```
```{r}
acc_s <- accessible_summary %>% group_by(dest) %>% summarize(ac_m = mean(total_time_mins), ac_s = sum(total_time_mins))
all_s <-all_summary %>% group_by(dest) %>% summarize(al_m = mean(total_time_mins), al_s = sum(total_time_mins))
left_join(acc_s, all_s, by = 'dest') %>% mutate(m_diff = ac_m - al_m, s_diff = ac_s - al_s, m_ratio = m_diff/al_m) %>% View
```
#### Visualize 90th percentile & median commutes to 137S (Chambers Street)
```{r}
dest <- "A31"
static_info_accessible <- get_heatmap_static_info(dest, graph_90, src = accessible_stations)
grid_info_accessible <- get_heatmap_grid_info(static_info_accessible, grid)

grid_info_accessible_manual <- get_heatmap_grid_info(static_info_accessible, grid, speed = 15)


# use the same color palette as above
static_info_full <- get_heatmap_static_info(dest, graph_90)
grid_info_full <- get_heatmap_grid_info(static_info_full, grid)

pal_full <- get_palette(grid_info_full)

map_full <- get_commute_heatmap(dest, static_info_full, grid_info_full, pal_full)
map_accessible <- get_commute_heatmap(dest, static_info_accessible, grid_info_accessible, pal_full)
map_accessible_manual <- get_commute_heatmap(dest, static_info_accessible, grid_info_accessible_manual, pal_full)
sync(map_accessible, map_full, map_accessible_manual)

```
```{r}

alternative_map <- function(map_data, palette = NULL, paletteSource = map_data){
  
  if(is.null(palette)){
    clrs <- brewer.pal(11, "RdYlGn")
  
    palette <- colorNumeric(
      palette = clrs,
      domain = sqrt(paletteSource$total_time_mins),
      reverse = TRUE,
      na.color = "transparent")
  }
  map_data <- arrange(map_data, desc(total_time_mins))
  
  leaflet(nyc_neighborhoods) %>%
    addTiles() %>%
    addProviderTiles("CartoDB.Positron") %>%
    setView(-73.98, 40.75, zoom = 10) %>%
    addLegend("bottomright", pal = palette,
              title = paste("commute time to", dest), value = sqrt(paletteSource$total_time_mins),
              labFormat = labelFormat(suffix = " mins", transform = function(x) x*x),
              opacity = 1) %>%
  addCircleMarkers(lng = map_data$lon, lat = map_data$lat, opacity = 1, fillOpacity = 1,
                   color =palette(sqrt(map_data$total_time_mins))) %>% 
    htmlwidgets::onRender(
    'function(el, x){
      overlay = el.querySelector(".leaflet-overlay-pane")
      overlay.style.opacity = ".6"
    }'
  )
}


giamf <- grid_info_accessible_manual %>% filter(total_time_mins < 120)
giff <- grid_info_full %>% filter(total_time_mins < 120)
clrs <- brewer.pal(11, "RdYlGn")
  
grid_pal <- colorNumeric(
  palette = clrs,
  domain = sqrt(giamf$total_time_mins),
  reverse = TRUE,
  na.color = "transparent")

sync(alternative_map(giamf, grid_pal),alternative_map(giff, grid_pal,giamf))
```

```{r}
grid_info_diff_time <- grid_info_accessible$total_time_mins - grid_info_full$total_time_mins
grid_info_diff <- select(grid_info_accessible, lat, lon)
grid_info_diff$total_time_mins <- grid_info_diff_time
grid_info_diff <-  grid_info_diff
alternative_map(grid_info_diff)
```
```{r}
grid_info_accessible_filtered <- grid_info_accessible %>% filter(walking_mins <= 20) 
grid_info_full_filtered <- grid_info_full %>% filter(walking_mins <= 20) 


clrs <- brewer.pal(11, "RdYlGn")
  
grid_pal <- colorNumeric(
  palette = clrs,
  domain = sqrt(grid_info_full_filtered$total_time_mins),
  reverse = TRUE,
  na.color = "transparent")

sync(alternative_map(grid_info_accessible_filtered, grid_pal, grid_info_full_filtered),alternative_map(grid_info_full_filtered, grid_pal))
```

```{r}
giafo <- grid_info_accessible %>% filter( total_time_mins < quantile(total_time_mins,.75))

giffo <- grid_info_full %>% filter(total_time_mins < quantile(total_time_mins,.75))


clrs <- brewer.pal(11, "RdYlGn")
  
grid_pal <- colorNumeric(
  palette = clrs,
  domain = sqrt(giffo$total_time_mins),
  reverse = TRUE,
  na.color = "transparent")

sync(alternative_map(giafo, grid_pal, giffo),alternative_map(giffo, grid_pal))
alternative_map(giafo, grid_pal, giffo)
```