---
title: "20180730_mta_triptime_distribution"
author: "Phoebe Nguyen, Akbar Mirza"
date: "July 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(ROCR)
library(here)
library(profvis)

source('itinerary_evaluation.R')
```
```{r load-data}
load(file = here("data", "todd_subway_realtime.RData"))
load(file= here("data", "line_color_dict.RData"))

```
```{r route color}
# routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
# routes$route_color <- replace_na(routes$route_color, "6D6E71")
# routes <- routes %>% mutate(route_color = sprintf("#%s", route_color))
#  
# line_color <- routes %>% group_by(route_color) %>% distinct(route_short_name, route_color) %>% arrange(route_color)
# hex_color <- data.frame(route_color= c(unique(line_color$route_color)),
#                           color_name= c("dark green","green", "blue","light green", "dark gray", "brown", "light gray", "purple", "red",  "yellow"  ,"orange" ))
#  
# route_color_dict <- left_join(line_color, hex_color)
# save(route_color_dict, file="line_color_dict.RData")


```

## Plotting Functions

```{r function-definitions}
compute_auc <- function(itin_pair_df) {
  # pull out the number from the ROCR package with pred <- predictions(predicted, actual); performance(pred, 'auc')@y.values[[1]]
  # or  with the pROC package: as.numeric(auc(response = actual, predictor = predicted))
  # itinerary is the actual value, trip time is the predicted value
  auc_df <- itin_pair_df %>%
    mutate(itin_line = as.factor(paste0(group, itin_id, sep=''))) %>%
    ungroup() %>%
    select(itin_line, time_diff)
  
  pred <- prediction(as.numeric(auc_df$time_diff, units = "mins"), as.factor(auc_df$itin_line))
  perf_nb <- performance(pred, 'auc')@y.values[[1]]
  perf <- performance(pred, 'auc')
  ifelse( perf_nb < 0.5, return(1 - perf_nb), return(perf))
}

# given a df with historical time_diffs for many initineraries
# will return a plot that maps avg time_diff vs variance
plot_frontier <- function(itins_df) {

  plot_data <- itins_df %>%
    #mutate(day_of_week = weekdays(leg1_start_time), hour = hour(leg1_start_time)) %>%
    # filter out weekends, and only get hour_start, hour_end
    #filter(isWeekend(day_of_week) == F, hour >= 7 & hour < 10) %>%
    group_by(itin_id, label) %>%
    mutate(avg = mean(time_diff), stdv = sd(time_diff))

  plot <- plot_data %>% ggplot(aes(x = stdv, y = avg)) +
    geom_point() +
    xlab('Standard Deviation of Trip Time') +
    ylab('Average Trip Time') +
    geom_text(aes(label = paste0(label, itin_id,sep='')), hjust = 0.5, vjust = 1)

  return(plot)
}
```

## Evaluating Interesting Multi-Leg Trips

Evaluating Trip Itineraries With Multiple Legs.

### Red Line vs Orange Line

Looking at trip time distributions between 86 St and 34 St on the Red Line (1, 2, 3) versus the Orange Line (B, D, F, M) 

#### Gather Red Line Data

```{r red-line-data}
red_line_data <- read_csv(here::here("data", "itineraries-Red-Line-86-34-train.csv"))

# filter out event and adding direction to station id
red_line_cleaned <- red_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
red_line_itins <- compute_all_itins(red_line_cleaned, subway_data)%>%
  mutate(group='red')
```

#### Gather Orange Line Data

```{r orange-line-data}
orange_line_data <- read_csv(here::here("data", "itineraries-Orange-Line-86-34-train.csv"))

# filter out event and adding direction to station id
orange_line_cleaned <- orange_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
orange_line_itins <- compute_all_itins(orange_line_cleaned, subway_data) %>%
  mutate(group='orange')

#combine first line df with second line df
red_orange_df <- bind_rows(red_line_itins, orange_line_itins)

# subway_line <- substr(unique(red_orange_df$label), 1, 1)[2] #this line picks off first character example(Q from trip label)
# subway_color <- route_color_dict[[which(route_color_dict$route_short_name == subway_line),         "route_color"]]
#   plot_colors <- c("#F7B731", subway_color)
  
```

#### Evaluating Differences Between Red & Orange Line & Plotting

```{r red-orange-comparison}
top_2_trips <- red_orange_df %>% group_by(group) %>% filter(itin_id == "1")
  
# Plot Distribution as Densities
plot_densities(top_2_trips,start_hour = 7 , end_hour = 10, bins = 60, route_color_dict)
ggsave(file = '../../figures/trip_time_red_express_vs_orange_local.png', width = 4, height = 4)
# Plot Frontiers
plot_frontier(red_orange_df)
ggsave(file = '../../figures/redX_vs_orgL_frontier_plot.png', width = 4, height = 4)
# Plot AUC for Top 2 Trips
compute_auc(top_2_trips)
```

### Midtown to Jamaica (JFK) during afternoon rush hour (3pm-5pm)

```{r midtown-jamaica-plots}

# load midtown-jamaica data, and clean for our uses
#   - filter out event and add direction to station id
mj_data <- read_csv(here::here("data", "itineraries-Midtown-Jamaica.csv")) %>%
  filter(event != "travel", itinerary_id < 2)

mj_df <- compute_all_itins(mj_data, subway_data, 15, 16) %>%
  filter(time_diff != expected_wait)

test_mj_df <- mj_df %>% filter(week(leg1_start_time) == 18:22)
# filtering out 0's from E train
plot_densities(mj_df, start_hour = 15 , end_hour = 17, bins = 20, route_color_dict)

# NOTE: DEBUG CODE
# subway_data %>% filter(stop_mta_id == "F12N" | stop_mta_id == "G06N", route_mta_id == "E") %>% group_by(realtime_trip_id) %>%
#   mutate(start_time = min(departure_time), count = n()) %>%
#   ungroup() %>% filter(realtime_trip_id %in% c(13391, 13404, 13457, 13495, 13517)) %>% View
```
### Midtown to Mets Willets (LGA) during afternoon rush hour (3pm-5pm)

```{r Bryant Park- Mets Willets}
bm_data <- read_csv(here::here("data","itineraries-5ave-MetsWillets.csv")) %>%
  filter(event != "travel", itinerary_id < 2)

bm_df <- compute_all_itins(bm_data, subway_data, 15,17) %>% filter(time_diff != expected_wait) # filtering out 0's from E train
plot_densities(bm_df, start_hour = 15 , end_hour = 17, bins = 60, route_color_dict)

mj_bm_comparison_df <- bind_rows(mj_df, bm_df)

plot_densities(mj_bm_comparison_df, start_hour = 15 , end_hour = 17, bins = 60, route_color_dict)
ggsave(file = '../../figures/trip_time_to_JFK_&_LGA.png', width = 4, height = 4)

```

### Red/Green to Fulton

#### Gathering Red Line Data from 96 - Fulton
```{r red-96-fulton-data}
red_line_data <- read_csv(here::here("data", "itineraries-96-Fulton-red.csv"))

# filter out event and adding direction to station id
red_line_cleaned <- red_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
red_line_itins <- compute_all_itins(red_line_cleaned, subway_data) %>%
  mutate(group='red')
```

#### Gathering Green Line Data from 96 - Fulton
```{r green-96-fulton-data}
green_line_data <- read_csv(here::here("data", "itineraries-96-Fulton-green.csv"))

# filter out event and adding direction to station id
green_line_cleaned <- green_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
green_line_itins <- compute_all_itins(green_line_cleaned, subway_data) %>%
  mutate(group='green')
```

#### Comparison of Red v. Green Data from 96 - Fulton
```{r red-green-96-fulton-comparison}
#combine first line df with second line df
red_green_df <- bind_rows(red_line_itins, green_line_itins)

red_green_top_pair <- red_green_df %>% group_by(group) %>% filter(itin_id == "1")
  
# Plot Distribution as Densities
plot_densities(red_green_top_pair, 7,10,100,route_color_dict)
ggsave(file = '../../figures/trip_time_red_vs_green_from_96_to_Fulton.png', width = 4, height = 4)

# Plot Frontiers
plot_frontier(red_green_df)
ggsave(file = '../../figures/frontier_plot_red_vs_green_from_96_to_Fulton.png', width = 4, height = 4)

# Plot AUC for Top 2 Trips
compute_auc(red_green_top_pair)
```

### Atlantic to Fulton - Super (Complex) Comparison

```{r atlantic-fulton-complex}
af_complex_data <- read_csv(here::here("data", "super-itin-atlantic-fulton.csv"))

# filter out event and adding direction to station id
af_complex_cleaned <- af_complex_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
af_complex_itins <- compute_all_itins(af_complex_cleaned, subway_data)

# plot distributions
plot_densities(af_complex_itins, 7,10,70,route_color_dict)
#ggsave(file = '../../figures/trip_times_from_Atlantic_Fulton.png', width = 4, height = 4)

plot_frontier(af_complex_itins, af_complex_itins$leg1_line)
# TODO: DELETE
# profvis({
#   #output for first data input
#   af_complex_itins.1 <- compute_all_itins(af_complex_cleaned, subway_data)
#   
#   #output for first data input
#   af_complex_itins.2 <- compute_all_itins(af_complex_cleaned, small_subway_data)
# })

```

### Atlantic to Union Square - Super (Complex) Comparison

```{r atlantic-union-sq-complex}
aus_complex_data <- read_csv(here::here("data", "super-itin-atlantic-union.csv"))

small_subway_data <- subway_data %>% filter(month(departure_time) == 2)
# filter out event and adding direction to station id
aus_complex_cleaned <- aus_complex_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
aus_complex_itins <- compute_all_itins(aus_complex_cleaned, subway_data)

# plot distributions
plot_densities(aus_complex_itins, 7,10, 100, route_color_dict)
```

### Atlantic to Bryant Park - Super (Complex) Comparison

```{r atlantic-bryant-complex}
abp_complex_data <- read_csv(here::here("data", "super-itin-atlantic-bryant.csv"))

# filter out event and adding direction to station id
abp_complex_cleaned <- abp_complex_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
abp_complex_itins <- compute_all_itins(abp_complex_cleaned, subway_data)

# plot distributions
plot_densities(abp_complex_itins, 7,10, 60, route_color_dict)
```

### Atlantic to Grand Central - Super (Complex) Comparison

```{r atlantic-grand-central-complex}
agc_complex_data <- read_csv(here::here("data", "super-itin-atlantic-grand-central.csv"))

# filter out event and adding direction to station id
agc_complex_cleaned <- agc_complex_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
agc_complex_itins <- compute_all_itins(agc_complex_cleaned, subway_data)

# plot distributions
plot_densities(agc_complex_itins, 7,10, 60, route_color_dict)
```

### Atlantic to Times Square - Super (Complex) Comparison

```{r atlantic-times-square-complex}
ats_complex_data <- read_csv(here::here("data", "super-itin-atlantic-times-square.csv"))

# filter out event and adding direction to station id
ats_complex_cleaned <- ats_complex_data %>%
  filter(event != "travel", itinerary_id < 3)

#output for first data input
ats_complex_itins <- compute_all_itins(ats_complex_cleaned, subway_data)

# plot distributions
plot_densities(ats_complex_itins, 7,10, 60,  route_color_dict)
```

## Plots for Debugging

```{r debugging_plots}

# TODO: Write A Function To Show Trains On Plot

trains = c("1", "2_3")
start_stop = c("120S", "123S")
end_stop = c("123S", "229S")
train_data <- vector("list", length(trains))

for (i in 1:length(trains)) {
  start <- start_stop[i]
  end <- end_stop[i]
  lines <- trains[i]
  
  leg_data <- get_leg_data(lines, start, end, subway_data)
  
  # NOTE: DEBUG CODE
  # print(nrow(leg_data))
  
  # append to list
  train_data[[i]] <- leg_data
}

# plot_data <- bind_rows(train_data)
# plot_data %>% View

debug_train_schedules_plot <- function(train_data, subway_data, start_hour = 8, end_hour = 9) {
  # filtered_df <- itins_df %>% filter(hour(leg1_start_time) >= start_hour & hour(leg1_start_time) <= end_hour)
  
  plot_data <- rbind_list(train_data)
  print(plot_data)
  plot_data %>%
    filter(month(start_time) == 2,
           day(start_time) == 16,
           hour(start_time) >= start_hour & hour(start_time) <= end_hour) %>%
    arrange(start_time) %>%
    mutate(itin = row_number()) %>%
    ggplot(aes(color = route_mta_id)) +
    geom_segment(aes(x = start_time, xend = departure_time, y = itin, yend = itin)) +
    geom_text(aes(label = realtime_trip_id, x = start_time, y = itin))
  
}

debug_train_schedules_plot(train_data, subway_data)

# NOTE: DEBUG PLOT - show all trains individually on plot

debug_itin_plot <- function(itin_df, subway_data, itin_number, num_trips = 5, start_hour = 8, end_hour = 9) {
  
  filtered <- itin_df %>% filter(itin_id == itin_number)
  
  reshaped_df <- filtered %>% select(matches("leg*")) %>% ungroup() %>%
    mutate(row_id = 1:n()) %>%
    gather(variable, value, -row_id) %>%
    extract(variable, c("leg", "var"), 'leg(.*?)_(.*)') %>%
    select(row_id, leg, var, value) %>%
    spread(var, value) %>%
    na.omit() %>%
    filter(row_id <= num_trips)
  
  reshaped_df$start_time <- as.POSIXct(as.numeric(reshaped_df$start_time), origin = "1970-01-01", tz = "UTC")
  reshaped_df$end_time <- as.POSIXct(as.numeric(reshaped_df$end_time), origin = "1970-01-01", tz = "UTC")
    
  
  plot <- reshaped_df %>% ggplot(aes(x = start_time, y = row_id)) +
    geom_segment(aes(xend = end_time, yend = row_id, color = as.factor(line))) +
    geom_point(aes(x = start_time, y = row_id)) +
    geom_point(aes(x = end_time, y = row_id)) +
    geom_text(aes(label = line), nudge_y = 0.25) 
  
  return(plot)
}

debug_itin_plot(green_line_itins, subway_data, 1)
plot <- debug_itin_plot(green_line_itins, subway_data, 1)
ggsave(plot = plot, file = here("figures", "trips-on-green-line-from-96-to-Fulton-between-7-and-8-am.png"), width = 8, height = 6, dpi = 300)

# filtered_test_df <- test_df %>% filter(hour(leg1_start_time) >= 8 & hour(leg1_start_time) <= 9)

# plot_data <- bind_rows(train_data[[1]], train_data[[2]], train_data[[3]])
# plot_data %>% filter(month(start_time) == 2 & day(start_time) == 16 & hour(start_time) > 19 & hour(start_time) < 24) %>%
#   arrange(start_time) %>%
#   mutate(itin = row_number()) %>%
#   ggplot(aes(color = route_mta_id)) +
#   geom_segment(aes(x = start_time, xend = departure_time, y = itin, yend = itin)) +
#   geom_text(aes(label = realtime_trip_id, x = start_time, y = itin))

# TODO: Write A Function To Show Itineraries On Plot

# ITINERARY PLOT
 red_green_top_pair %>% mutate(itin = row_number()) %>%
   filter(itin <= 5) %>%
   ggplot() +
     geom_segment(aes(x = leg1_start_time, xend = leg1_end_time, y = itin, yend = itin), color = 'red') +
     geom_segment(aes(x = leg2_start_time, xend = leg2_end_time, y = itin, yend = itin), color = 'blue') +
     geom_segment(aes(x = leg3_start_time, xend = leg3_end_time, y = itin, yend = itin), color = 'green') +
     geom_text(aes(label = leg1_trip_id, x = leg1_start_time, y = itin), color = "red") +
     geom_text(aes(label = leg2_trip_id, x = leg2_start_time, y = itin), color = "blue") +
     geom_text(aes(label = leg3_trip_id, x = leg3_start_time, y = itin), color = "green") +
     ylim(c(0, 10))

# filtered_test_df %>% ggplot() +
#   geom_histogram(aes(x = time_diff))
```