---
title: "20180717_mta_triptime_distribution"
author: "Phoebe Nguyen, Akbar Mirza"
date: "July 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(here)

# load todd's subway_data
load(file = "../../data/todd_subway_realtime.RData")
# NOTE: How We Get Subway_data_2
# subway_data.2 <- subway_data %>%
#   group_by(realtime_trip_id) %>%
#   arrange(desc(departure_time)) %>%
#   mutate(arrival_time=lead(departure_time)) %>%
#   ungroup()
# 
# save(subway_data.2, file = "../../data/todd_subway_realtime_2.RData")
load(file = "../../data/todd_subway_realtime_2.RData")



# load stop_times from GTFS Static
stop_times <- read_csv("../../data/google_transit_subway_static/stop_times.txt")

# load station_data from mta
station_data <- read_csv("http://web.mta.info/developers/data/nyct/subway/Stations.csv")

# get stop_id and stop_name fields, and create a stop_id with N and S appended
station_data <- station_data %>% group_by(`GTFS Stop ID`, `Stop Name`) %>%
  select(stop_id = `GTFS Stop ID`, stop_name = `Stop Name`) %>%
  mutate(stop_id_N = sprintf('%sN', stop_id), stop_id_S = sprintf('%sS', stop_id)) %>%
  gather(`stop_id_N`, `stop_id_S`, key = "stop_ids", value = "stop_id") %>%
  select(stop_name, stop_id)

################################################################################
# Getting Station Information
################################################################################
# Generate edges between stations
trips <- stop_times %>% extract(trip_id, "route", regex=".*_.*_(.*)\\.\\..*", remove = FALSE)
trip_edges <- trips %>% mutate(prev_stop_id = ifelse(lag(trip_id) == trip_id, lag(stop_id), NA))
edges <- trip_edges %>% select(route, stop_id, prev_stop_id) %>% distinct()
# create stations dataframe with line, stop_id, stop_name
stations <- edges %>% left_join(station_data, by = c("stop_id"="stop_id")) %>% select(line=route, stop_id, stop_name)
```

## Express vs Local Train Distribution

- Filter through all local and express train trips between two stations. 
- Calculate the time between two stations (trip time) for all trip across the day
- Compare the differences in histogram

```{r express-v-local-dist}
# Helper Function to Determine if Day is Weekend
isWeekend <- function(day_of_week) {
  ifelse(day_of_week == "Saturday" | day_of_week == "Sunday", T, F)
}

stops <- stations %>%
  filter(line == "1", grepl(sprintf("%s$", "S"), stop_id)) %>%
  unique()

# find trip time during 96th station and 14th station

plot_local_express <- function(local_line, express_line, start_stop_id, end_stop_id)
{
  
  
  # local train
  local_train <- subway_data %>%
    filter(stop_mta_id == start_stop_id | stop_mta_id == end_stop_id,
           route_mta_id == local_line) %>%
    mutate(day_of_week = weekdays((departure_time)),
           hour = hour(departure_time)) %>%
    filter(isWeekend(day_of_week) == F, hour > 6, hour < 20) %>%
    left_join(stops, by = c("stop_mta_id" = "stop_id")) %>%
    group_by(realtime_trip_id) %>% 
    arrange(desc(departure_time)) %>%
    mutate(diff = (first(departure_time) - last(departure_time))/60)
  
  # plot for local line
  plot_local <- local_train %>%
    ggplot(aes(x=diff)) + geom_histogram()

  # express train
  express_train <- subway_data %>%
    filter(stop_mta_id == start_stop_id | stop_mta_id == end_stop_id,
           route_mta_id == express_line) %>%
    mutate(day_of_week = weekdays(departure_time),
           hour = hour(departure_time)) %>%
    filter(isWeekend(day_of_week) == F, hour > 6, hour < 20) %>%
    left_join(stops, by = c("stop_mta_id" = "stop_id")) %>%
    group_by(realtime_trip_id) %>%
    arrange(desc(departure_time)) %>%
    mutate(diff = (first(departure_time) - last(departure_time))/60)
  
  # plot for express train
  plot_express <- express_train %>%
    ggplot(aes(x=diff)) + geom_histogram()

  # code for combining plots as per: https://stackoverflow.com/questions/21192002/how-to-combine-2-plots-ggplot-into-one-plot
  local_train$group <- "local"
  express_train$group <- "express"
  
  combined <- rbind(local_train, express_train)
  
  combined_plot <- ggplot(combined, aes(x=diff, group=group, col=group, fill=group)) +
    geom_histogram(alpha = 0.8, position = "identity")
  # NOTE: can get intersecting distributions by running the code below 
  # geom_density(alpha = 0.8, position = "identity")
  
  return(combined_plot)
}

plot_local_express ("1", "2", "120S", "132S")

plot <- plot_local_express ("1", "2", "120S", "132S")
#ggsave(plot = plot, file = here("figures", "local-vs-express-96-14-St.png"), width = 5, height = 4, dpi = 300)
```

## Introducing One Stop Transfer

- Explore a slightly different route on this line that could use either the local or a combination of the local and express with a transfer in between(e.g., 18th to 96th st.)
- Compare the distribution of trip times using the local to those using the local + express

```{r compute-trip-time}

# NOTE: DEBUG CODE  
start_station <- "120S"
start_line <- "2"
transfer_station <- "127S"
transfer_line <- "1"
end_station <- "131S"
start_time <- "2018-01-20 09:30:00"

compute_trip_time <- function (start_station, start_line, transfer_station, transfer_line, end_station, start_time) {
    
    # find first trips on the start_line at the start_station that depart after start_time
    first_trains <- subway_data.2 %>%
      filter(stop_mta_id == start_station | stop_mta_id == transfer_station,
             route_mta_id == start_line) %>%
      filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
      group_by(realtime_trip_id) %>%
      mutate(count = n()) %>%
      filter(count >= 2) %>%
      ungroup() %>%
      arrange(departure_time)
    # ungroup and arrange by departure_time
    
    # get the trip_id of the first trip after the depart_time
    first_trip_id <- first(first_trains$realtime_trip_id)
    
    # NOTE: arrival_time to first_train # 971896
    
    # find the arrival time for trip_id at transfer station
    transfer_arrival_time <- subway_data.2 %>%
      filter(realtime_trip_id == first_trip_id,
             stop_mta_id == transfer_station) %>%
      .$arrival_time
    
    # find the first trip on the transfer_line at the transfer_station that departs after the arrival_time of our express train
    transfer_trains <- subway_data.2 %>%
      filter(stop_mta_id == transfer_station | stop_mta_id == end_station,
             route_mta_id == transfer_line) %>%
      filter(ymd_hms(departure_time) > ymd_hms(transfer_arrival_time)) %>%
      group_by(realtime_trip_id) %>%
      mutate(count = n()) %>%
      filter(count >= 2) %>%
      ungroup() %>%
      arrange(departure_time)
      # note above
    
    transfer_trip_id <- first(transfer_trains$realtime_trip_id)
    
    # NOTE: arrival_time to transfer_train
    
    # compute the time difference of when our train arrives at end_station and start_time
    dest_arrival_time <- subway_data.2 %>%
      filter(realtime_trip_id == transfer_trip_id,
             stop_mta_id == end_station) %>%
      .$arrival_time
    
    time_diff <- as.numeric(difftime(ymd_hms(dest_arrival_time), ymd_hms(start_time)), units = "mins")
    
     return(time_diff)
#     
}
# 
compute_trip_time("120S", "2", "127S", "1", "131S", "2018-01-22 09:30:00")

```

## Comparing Distributions of Local vs Local + Express

### Computing Local + Express Wait Times On Large Scale Efficiently

```{r example-for-trips}

local_line <- "1"
express_line <- "2"
start_stop_id <- "120S"
transfer_stop_id <- "127S"
end_stop_id <- "131S"
start_time <- "2018-01-20 09:30:00"

# Getting 2 Trains
data_2_train <- subway_data.2 %>%
  filter(stop_mta_id == "120S" | stop_mta_id == "127S",
         route_mta_id == "2") %>%
  # filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
  group_by(realtime_trip_id) %>%
  mutate(start_time = min(departure_time),
         day_of_week = weekdays((departure_time)),
           hour = hour(departure_time)) %>%
  filter(n() >= 2, stop_mta_id == "127S",isWeekend(day_of_week) == F, 
         hour > 6, hour < 20) %>%
  arrange(arrival_time)

# Getting 1 Trains
data_1_train <- subway_data.2 %>%
  filter(stop_mta_id == "127S" | stop_mta_id == end_stop_id,
         route_mta_id == local_line) %>%
  # filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
  group_by(realtime_trip_id) %>%
  mutate(end_time = max(departure_time),
         day_of_week = weekdays((departure_time)),
           hour = hour(departure_time)) %>%
  filter(n() >= 2, 
         stop_mta_id == "127S",
         isWeekend(day_of_week) == F, 
         hour > 6, hour < 20) %>%
  arrange(departure_time)



# NOTE: train 1 start time and train 2 end time for debugging and evaluating purposes
i = 1
j = 1
times <- vector('numeric')
start_times <- vector('double')
transfer_arrival_times <- vector('double')
end_times <- vector('double')
transfer_departure_times <- vector('double')
realtime_start_ids <- vector('numeric')
realtime_end_ids <- vector('numeric')

while (i < nrow(data_2_train)) {
  if (data_1_train$departure_time[j] > data_2_train$arrival_time[i]) {
    # df <- data.frame(trip_time = 0, start_trip = 0, end_trip = 0)
    realtime_start_ids <- c(realtime_start_ids, data_2_train$realtime_trip_id[i])
    start_times <- c(start_times, data_2_train$start_time[i])
    transfer_arrival_times <- c(transfer_arrival_times, data_2_train$arrival_time[i])
    realtime_end_ids <- c(realtime_end_ids, data_1_train$realtime_trip_id[j])
    transfer_departure_times <- c(transfer_departure_times, data_1_train$departure_time[j])
    end_times <- c(end_times, data_1_train$end_time[j])
    times <- c(times, as.numeric(difftime(data_1_train$end_time[j], data_2_train$start_time[i]), units = "mins"))
    i = i + 1
  } else {
    j = j + 1
  }
}

# verify with real data
new_df <- data.frame(start_trip_id = realtime_start_ids,
                     starts = as.POSIXct(start_times, origin = "1970-01-01", tz = "UTC"),
                     transfer_starts = as.POSIXct(transfer_arrival_times, origin = "1970-01-01", tz = "UTC"),
                     end_trip_id = realtime_end_ids,
                     transfer_ends = as.POSIXct(transfer_departure_times, origin = "1970-01-01", tz = "UTC"),
                     ends = as.POSIXct(end_times, origin = "1970-01-01", tz = "UTC"),
                     diffs = times) 
# Views to Verify Trips
# ORIGINAL SUBWAY DATA
#subway_data %>% filter(realtime_trip_id == 3300 & stop_mta_id == "120S" | realtime_trip_id == 3327 & stop_mta_id == "131S") %>% View("subway_data 3300")
# NEW DATA FRAME DATA
#new_df %>% filter(start_trip_id == 3300) %>% View("new_df 3300")


# TODO: figure out multiple transfers with multiple lists
```

### Plotting Distributions of Local vs Local + Express

```{r plotting-dist-local-v-local+express}

# Local Train Data & Distribution
local_line <- "1"
express_line <- "2"
start_stop_id <- "120S"
end_stop_id <- "131S"

# local train
local_diffs <- subway_data %>%
  filter(stop_mta_id == start_stop_id | stop_mta_id == end_stop_id,
         route_mta_id == local_line) %>%
  mutate(day_of_week = weekdays((departure_time)),
           hour = hour(departure_time)) %>%
  filter(isWeekend(day_of_week) == F, hour > 6, hour < 20) %>%
  left_join(stops, by = c("stop_mta_id" = "stop_id")) %>%
  group_by(realtime_trip_id) %>% 
  arrange(desc(departure_time)) %>%
  mutate(diff = as.numeric(difftime(first(departure_time),last(departure_time)), units = "mins")) %>%
  select(realtime_trip_id, diff) %>%
  unique() %>% ungroup() %>%
  select(diffs=diff)


# plot for local line
plot_local <- local_diffs %>%
  ggplot(aes(x=diffs)) + geom_histogram()

# express train
local_with_express_diffs <- new_df %>%
  select(diffs)

# plot for express train
plot_express <- local_with_express_diffs %>%
  ggplot(aes(x=diffs)) + geom_histogram()

# code for combining plots as per: https://stackoverflow.com/questions/21192002/how-to-combine-2-plots-ggplot-into-one-plot
local_diffs$group <- "local"
local_with_express_diffs$group <- "express + local"

combined <- rbind(local_diffs, local_with_express_diffs)

combined_plot <- ggplot(combined, aes(x=diffs, group=group, col=group, fill=group)) +
  geom_histogram(alpha = 0.8, position = "identity") +
  labs(title = "Trip Time Evaluation", x = "Trip Time (in minutes)", y = "Number of Trips") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.key.size = unit(1.5, 'lines')) +
  xlim(c(0, 50))
# NOTE: can get intersecting distributions by running the code below 
# geom_density(alpha = 0.8, position = "identity")

combined_plot
ggsave(plot = combined_plot, file = here("figures", "local-vs-express&local-96-18-St-during-work-hour.png"), width = 5, height = 4, dpi = 300)

```