---
title: "Map With K Shortest Paths"
author: "Graph Team"
date: "July 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Sasha, Ayliana, Brian, & Taxi*  

#### Load libraries
```{r libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggmap)
library(leaflet)
library(igraph)
library(scales)
library(lubridate)
```


#### Load MTA data & Google Transit static files
```{r load_data, warning=FALSE, message=FALSE}
stations <- read_csv('http://web.mta.info/developers/data/nyct/subway/Stations.csv')
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
trips <- read_csv('../../data/google_transit_subway_static/trips.txt')
stop_times <- read_csv('../../data/google_transit_subway_static/stop_times.txt')
transfers <- read_csv('../../data/google_transit_subway_static/transfers.txt')

load('../../data/todd_subway_realtime.RData')
```


#### The `stations` data looks like:
```{r stations}
as.data.frame(stations) %>% head
```


#### We might want to view each route on its own line
```{r warning=FALSE}
new_stations <- stations %>% 
  select(complex_id = `Complex ID`, stop_id = `GTFS Stop ID`, stop_name = `Stop Name`, 
         stop_lat = `GTFS Latitude`, stop_lon = `GTFS Longitude`, line=Line, daytime_routes=`Daytime Routes`)

new_stations <- new_stations %>% 
  separate(daytime_routes, into=c('route1', 'route2', 'route3', 'route4'), sep=" ") 

ss1 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route1)
ss2 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route2)
ss3 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route3)
ss4 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route4)

routes_by_lines <- rbind(ss1, ss2, ss3, ss4) %>% filter(!is.na(route))

as.data.frame(routes_by_lines) %>% head
```


#### Use `stop_times` to understand the schedules/sequences of each line
```{r}
connections <- stop_times %>%
  left_join(stops) %>% 
  extract(trip_id, c("route_id"), regex=".*_.*_(.*)\\.\\..*", remove=FALSE) %>%
  mutate(stop_id = substr(stop_id, 1, 3),
    prev_stop_id = ifelse(trip_id == lag(trip_id), lag(stop_id), NA),
    prev_stop_lat = ifelse(trip_id == lag(trip_id), lag(stop_lat), NA),
    prev_stop_lon = ifelse(trip_id == lag(trip_id), lag(stop_lon), NA),
    prev_stop_name = ifelse(trip_id == lag(trip_id), lag(stop_name), NA))

sequences <- connections %>%
  left_join(routes) 

# Catch NA color values
sequences$route_color <- replace_na(sequences$route_color, "000000") 

sequences <- sequences %>%
  mutate(route_color = sprintf("#%s", route_color)) %>%
  select(route_id, stop_id, stop_name, stop_lat, stop_lon, 
         prev_stop_id, prev_stop_name, prev_stop_lat, prev_stop_lon,
         route_color) %>%
  distinct 

```


#### Use `transfers` to understand where we can transfer to other lines
```{r transfers}

# Prepping the transfer data
transfer_sequences <- transfers %>% left_join(stops, by = c("to_stop_id" = "stop_id")) %>% 
  left_join(stops, by = c("from_stop_id" = "stop_id")) %>%
  mutate(route_ids = "T", route_color = "#000000") %>%
  select(route_ids, stop_id = to_stop_id, stop_name = stop_name.x, stop_lat = stop_lat.x, stop_lon = stop_lon.x, prev_stop_id = from_stop_id, prev_stop_name = stop_name.y, prev_stop_lat = stop_lat.y, prev_stop_lon = stop_lon.y, route_color, weight = min_transfer_time)

```


#### Get weights for the connections between stations
```{r weights, warning=FALSE}

# Compute travel time for a trip between given stations and take the average of the travel time
station_weights <- realtime %>%
  mutate(stop_mta_id = substr(stop_mta_id, 1, 3)) %>%
  arrange(realtime_trip_id, departure_time) %>% 
  mutate(travel_time = ifelse(realtime_trip_id == lag(realtime_trip_id), departure_time - lag(departure_time), NA)) %>%
  mutate(prev_stop_mta_id = ifelse(realtime_trip_id == lag(realtime_trip_id), lag(stop_mta_id), NA)) %>%
  filter(!is.na(travel_time)) %>% 
  group_by(stop_mta_id, prev_stop_mta_id) %>% 
  summarize(weight = mean(travel_time)) 

station_route_ids <- sequences %>% select(route_id, stop_id, prev_stop_id) %>% distinct() %>% filter(! is.na(prev_stop_id)) %>%
  spread(key = route_id, value = 1) %>% 
  unite("route_ids",-stop_id, -prev_stop_id) %>%
  mutate(route_ids = gsub("(_NA)|(NA_)", "", route_ids))

# Join with the sequences dataframe to get back information about the stops
full_sequences <- sequences %>% left_join(station_weights, by = c("stop_id" = "stop_mta_id", "prev_stop_id" = "prev_stop_mta_id")) %>% 
  select(-route_id) %>% distinct() %>% 
  filter(! is.na(prev_stop_id)) %>% 
  left_join(station_route_ids) 
  


# Transfers already have weights from the min_transfer_time field (now renamed weight) so we rbind the two sequence objects
full_sequences <- rbind(full_sequences, transfer_sequences)

```


#### Get better weights for the connections between stations
```{r day_time_weights}

# filter_df <- function(day, time) {
#  time <- as.POSIXct(time, format="%H:%M:%S")
#  filtered_weights <- realtime %>%
#    mutate(lagged_departure_time = lag(departure_time),
#           day_of_week = weekdays(departure_time)) %>%
#    filter(day_of_week == day, hour(departure_time) >= hour(time)-1, hour(departure_time) <= hour(time)+1) %>%
#    mutate(stop_mta_id = substr(stop_mta_id, 1, 3)) %>%
#    arrange(realtime_trip_id, departure_time) %>% 
#    mutate(travel_time = ifelse(realtime_trip_id == lag(realtime_trip_id), departure_time - lag(departure_time), NA)) %>%
#    mutate(prev_stop_mta_id = ifelse(realtime_trip_id == lag(realtime_trip_id), lag(stop_mta_id), NA)) %>%
#    arrange(realtime_trip_id, departure_time) %>%
#    filter(!is.na(travel_time)) %>% 
#    group_by(route_mta_id, stop_mta_id, prev_stop_mta_id) %>% 
#    summarize(weight = mean(travel_time))
#  
#  filtered_weights$prev_stop_mta_id <- as.character(filtered_weights$prev_stop_mta_id)
# 
#   full_sequences <- sequences %>%
#    left_join(filtered_weights, by = c("stop_id" = "stop_mta_id", "prev_stop_id" = "prev_stop_mta_id", "route_id" = "route_mta_id"))
# 
#  return(full_sequences)
# }
# 
# 
# get_weighted_map_igraph <- function(df, day, time){
#  
#  # historical data for weights
#  weights_df <- filter_df(day, time) %>%
#    select(route_id, stop_id, prev_stop_id, weight)
# 
#  igraph_edges <- weights_df %>%
#    filter(!is.na(prev_stop_id), !is.na(weight)) %>%
#    select(prev_stop_id, stop_id, weight)
# 
#  mta_igraph <- graph.data.frame(igraph_edges, directed=TRUE)
#  plot(mta_igraph)
#  
# }
# 
#get_weighted_map_igraph(map_data, "Weekday", "6:30:00")

```

<!-- #### Make a subway map! -->
<!-- ```{r ggmap, warning=FALSE} -->

<!-- nyc_map <- get_map(location = c(lon = -73.9568247, lat = 40.7202688), maptype = "terrain", zoom = 12) -->
<!-- ggmap(nyc_map) + -->
<!--  geom_point(data = full_sequences, aes(x = stop_lon, y = stop_lat)) + -->
<!--  geom_segment(data = full_sequences, aes(x=prev_stop_lon, y=prev_stop_lat, xend=stop_lon, yend=stop_lat, color = route_id)) -->

<!-- ``` -->


<!-- #### Make another (prettier) map -->
<!-- ```{r leaflet} -->

<!-- map <- leaflet() %>% -->
<!--   addTiles() %>% -->
<!--   setView(-74.00, 40.71, zoom = 12) %>% -->
<!--   addProviderTiles("CartoDB.Positron") %>% -->
<!--   addCircleMarkers(sequences$stop_lon, sequences$stop_lat, color = sequences$route_color, -->
<!--                    popup = full_sequences$stop_name, -->
<!--                    radius = 3) -->

<!-- for (i in 1:nrow(full_sequences)) { -->
<!--   map <- map %>% -->
<!--     addPolylines(lat = c(full_sequences[i,]$stop_lat, full_sequences[i,]$prev_stop_lat), -->
<!--                  lng = c(full_sequences[i,]$stop_lon, full_sequences[i,]$prev_stop_lon), -->
<!--                  color = full_sequences[i,]$route_color, -->
<!--                  dashArray = ifelse(full_sequences[i,]$route_id == "T", c(8), c(0)), -->
<!--                  weight = 1) -->
<!-- } -->
<!-- map -->

<!-- ``` -->

#### Make an igraph
```{r igraph}

theme_set(theme_bw())

# Input a dataframe of the edges that exist between stops, with weight as an attribute
igraph_edges <- full_sequences %>% filter(!is.na(prev_stop_id), !is.na(weight)) %>% select(prev_stop_id, stop_id, weight, route_ids) %>% distinct()
mta_igraph <- graph.data.frame(igraph_edges, directed=TRUE)

plot(mta_igraph)

```

#### Find shortest route
```{r shortest_route}

# Get the shortest path between two stops
shortest_path <- shortest_paths(mta_igraph, from = c("701"), to = c("R19"), mode="out", output="both")

# Making a tidy dataframe out of the generated path
shortest_route <- data.frame(stop_id = (sapply(shortest_path$vpath, as_ids)), stringsAsFactors = FALSE) %>% left_join(stops, by = "stop_id") %>% select(stop_id, stop_name)

shortest_route

```

#### Function to find shortest k paths
```{r k_shortest_function}

##################
# Helper Functions
##################

# Determine if a given path variant has already been seen
contains.path <- function(variants, variant){
  return( any( unlist( lapply( variants, function(x){ identical(unlist(x$variant$vert),unlist(variant)) } ) ) ) )
}

# Choose shortest path variant
select.shortest.path <- function(variants){
  return( which.min( unlist( lapply( variants, function(x){x$variants$dist} ) ) ) )
}

# Return all variants of a given path
calculate.variants <- function(variants, variant, from, to, gmode){
  # Take graph from current path
  g <- variant$g
  
  # Iterate through edges, removing one each iterations
  for (j in unlist(variant$path)){
    newgraph <- delete.edges(g, j) # Remove edge
    sp <- get.shortest.paths(newgraph,from,to, output='both', mode=gmode) # Calculate shortest path
    spd <- shortest.paths(newgraph,from,to,mode=gmode) # Calculate length
    if (spd != Inf){ # The the path is found
      if (!contains.path(variants, sp$vpath)) # Add to list, unless it already contains the same path
      {
        variants[[length(variants)+1]] <- list(g=newgraph, variants=list(path=sp$epath, vert=sp$vpath, dist=spd))
      }
    }
  }
  return(variants)
}

######################
# Main K Path Function
######################

k.shortest.paths <- function(graph, from, to, k, gmode){
  # First shortest path
  k0 <- get.shortest.paths(graph,from,to, output='both', mode=gmode)
  
  # Number of currently found shortest paths
  kk <- 1
  
  # List of alternatives
  variants <- list()
  
  # Shortest variants
  shortest.variants <- list(list(g=graph, path=k0$epath, vert=k0$vpath, dist=shortest.paths(graph,from,to,mode=gmode)))
  
  # Until k shortest paths are found
  while(kk<k){
    # Take last found shortest path
    last.variant <- shortest.variants[[length(shortest.variants)]]              
        
    # Calculate all alternatives
    variants <- calculate.variants(variants, last.variant, from, to, gmode)
        
    # Find shortest alternative
    sp <- select.shortest.path(variants)
    
    # Add to list, increase kk, remove shortest path from list of alternatives
    shortest.variants[[length(shortest.variants)+1]] <- list(g=variants[[sp]]$g, path=variants[[sp]]$variants$path, vert=variants[[sp]]$variants$vert, dist=variants[[sp]]$variants$dist)
    kk <- kk+1
    variants <- variants[-sp]
  }
  return(shortest.variants)
}

```


#### Find k shortest paths
```{r k_shortest, warning=FALSE}

k_num <- 2

#Find k_num shortest paths leading going from stop 701 to stop R19 
k_shortest <- k.shortest.paths(mta_igraph, from = "701", to = "R19", k=k_num, gmode="out")

shortest_routes <- setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("itinerary_id", "line", "station","direction", "event"))

for(i in 1:k_num){
  station_list <- names(unlist(k_shortest[[i]]$`vert`))
  line_ids <- edge.attributes(mta_igraph, index = unlist(k_shortest[[i]]$`path`))$route_ids
  route = vector(mode = "character")
  
  prev_line_ids <- line_ids[1]
  for (j in 1:length(station_list)){
    temp <- data.frame(itinerary_id = i, line = line_ids[j], station =
                         station_list[j], direction = "N", event = NA)
    
    shortest_routes <- rbind(shortest_routes, temp)
    
    #cur_line_ids <- str_split(line_ids[j], "_")
    #intersect <- as.list(intersect(cur_line_ids, prev_line_ids))
    #if (length(intersect) == 0){
    #  append(route, rep(prev_line_ids), j-length(route))
     # intersect <- line_ids[j+1]
    #}
   # prev_line_ids <- intersect
  }
  
 # shortest_routes$line <- route
}

```