---
title: "Graph/Itinerary"
author: "Sasha, Ayliana, Taxi"
date: "July 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Sasha, Ayliana, Brian, & Taxi*  

#### Load libraries
```{r}
library(tidyverse)
library(igraph)
library(lubridate)
```

#### Load data
```{r}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
trips <- read_csv('../../data/google_transit_subway_static/trips.txt')
stop_times <- read_csv('../../data/google_transit_subway_static/stop_times.txt')
transfers <- read_csv('../../data/google_transit_subway_static/transfers.txt')

load('../../data/todd_subway_realtime.RData')

```

#### Set filters for time intervals/day
```{r}

# all service
# time_filter <- seq(0, 23)
# day_filter <- c("Weekday", "Saturday", "Sunday")

# ordinary (weekday+daytime) service
time_filter <- seq(6, 23)
day_filter <- c("Weekday")

# weekend (daytime) service
#time_filter <- seq(6, 23)
#day_filter <- c("Saturday", "Sunday")

# late night (weekday+weekend) service
# time_filter <- seq(0, 5)
# day_filter <- c("Weekday", "Saturday", "Sunday")
```

#### Formatting static
```{r}
trips$route_id <- as.character(trips$route_id)


# all scheduled trips
all_trips <- stop_times %>%
  filter(!is.na(arrival_time)) %>%
  left_join(stops) %>%
  extract(trip_id, c("route_id"), regex=".*_.*_([^.]*)\\.\\.?.*", remove=FALSE) %>%
  extract(trip_id, c("day_of_week"), regex=".*-.*-(.*)-.*", remove=FALSE) %>%
  extract(trip_id, c("time"), regex=".*-.*-.*-.*_(.*)_.*\\.\\.?.*", remove=FALSE) %>%
  mutate(direction = substr(stop_id, 4, 4),
         stop_id = substr(stop_id, 1, 3),
         prev_stop_id = ifelse(trip_id == lag(trip_id), lag(stop_id), NA),
         prev_stop_name = ifelse(trip_id == lag(trip_id), lag(stop_name), NA),
         trip_start_time = seconds_to_period(as.numeric(time)*.6),
         trip_start_time = as.POSIXct(sprintf("%s:%s:%s", 
                                              hour(trip_start_time), minute(trip_start_time), second(trip_start_time)),
                                      "%H:%M:%S", tz="America/New_York")) %>%
  left_join(trips) %>%
  select(route_id, trip_id, direction_id = direction, day_of_week, trip_start_time, arrival_time, departure_time, 
         stop_id, stop_name, prev_stop_id, prev_stop_name)

```

#### Transfer formatting
```{r}
# prepping the transfer data
transfer_sequences <- transfers %>% left_join(stops, by = c("to_stop_id" = "stop_id")) %>% 
  left_join(stops, by = c("from_stop_id" = "stop_id")) %>%
  mutate(route_ids = "T", direction_id = NA, weight = min_transfer_time, sd = NA, lower_quartile = NA, median = NA, upper_quartile = NA) %>%
  select(route_ids, direction_id, stop_id = to_stop_id, stop_name = stop_name.x, prev_stop_id = from_stop_id, 
         prev_stop_name = stop_name.y, weight, sd, lower_quartile, median, upper_quartile)
```

#### Realtime Formatting
```{r}
realtime <- realtime %>%
  mutate(day_of_week = weekdays(departure_time),
         day_of_week = ifelse(day_of_week != "Saturday" & day_of_week != "Sunday",
                              "Weekday", day_of_week),
         direction_id = ifelse(direction == 1, "N", "S"))
```


### RERUN FROM THIS POINT ON IF TIME/DAY FILTER CHANGES
#### Apply time/day filter here
```{r}
# unique trip sequences by line
unique_sequences <- all_trips %>%
  filter(hour(trip_start_time) %in% time_filter) %>%
  filter(day_of_week %in% day_filter) %>%
  select(route_id, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name) %>%
  distinct
```


#### Get weights for the connections between stations
```{r weights, warning=FALSE}

station_weights <- realtime %>%
  mutate(stop_mta_id = substr(stop_mta_id, 1, 3)) %>%
  arrange(realtime_trip_id, departure_time) %>% 
  mutate(travel_time = ifelse(realtime_trip_id == lag(realtime_trip_id),
                              departure_time - lag(departure_time), NA)) %>%
  mutate(prev_stop_mta_id = ifelse(realtime_trip_id == lag(realtime_trip_id),
                                   lag(stop_mta_id), NA)) %>%
  filter(!is.na(travel_time), 
         hour(departure_time) %in% time_filter,
         day_of_week %in% day_filter) %>% 
  group_by(route_mta_id, stop_mta_id, prev_stop_mta_id) %>% 
  summarize(weight = mean(travel_time), sd = sd(travel_time, na.rm=TRUE),
            lower_quartile = quantile(travel_time, 0.25),
            median = median(travel_time), upper_quartile = quantile(travel_time, 0.75))
```

#### Combine station ids
```{r}
station_route_ids <- unique_sequences %>% select(route_id, stop_id, prev_stop_id) %>% distinct() %>% filter(! is.na(prev_stop_id)) %>%
  spread(key = route_id, value = 1) %>% 
  unite("route_ids",-stop_id, -prev_stop_id) %>%
  mutate(route_ids = gsub("(_NA)|(NA_)", "", route_ids))
```

#### Combine realtime with static edges
```{r}
# Static with realtime weights and route_ids
full_sequences <- unique_sequences %>%
  left_join(station_weights, by = c("stop_id" = "stop_mta_id", "prev_stop_id" =
                              "prev_stop_mta_id", "route_id" = "route_mta_id"))%>%
  left_join(station_route_ids, by = c("stop_id", "prev_stop_id")) %>%
  select(route_ids, direction_id, stop_id, stop_name, prev_stop_id, prev_stop_name, 
         weight, sd, lower_quartile, median, upper_quartile)
  
# Transfers already has weights from the min_transfer_time field (now renamed weight) so we rbind the two sequence objects
full_sequences <- rbind(full_sequences, transfer_sequences)

```

#### Make an igraph
```{r igraph}

theme_set(theme_bw())

# Input a dataframe of the edges that exist between stops, with weight and direction as attributes
igraph_edges <- full_sequences %>%
  filter(!is.na(prev_stop_id), !is.na(weight)) %>%
  select(prev_stop_id, stop_id, weight, route_ids, direction_id)

mta_igraph <- graph.data.frame(igraph_edges, directed=TRUE)

```

#### Function to find shortest k paths
```{r k_shortest_function}

##################
# Helper Functions
##################

# Determine if a given path variant has already been seen
contains.path <- function(variants, variant){
  return( any( unlist( lapply( variants, function(x){ identical(unlist(x$variant$vert),unlist(variant)) } ) ) ) )
}

# Choose shortest path variant
select.shortest.path <- function(variants){
  return( which.min( unlist( lapply( variants, function(x){x$variants$dist} ) ) ) )
}

# Return all variants of a given path
calculate.variants <- function(variants, variant, from, to, gmode){
  # Take graph from current path
  g <- variant$g
  
  # Iterate through edges, removing one each iterations
  for (j in unlist(variant$path)){
    newgraph <- delete.edges(g, j) # Remove edge
    sp <- get.shortest.paths(newgraph,from,to, output='both', mode=gmode) # Calculate shortest path
    spd <- shortest.paths(newgraph,from,to,mode=gmode) # Calculate length
    if (spd != Inf){ # The the path is found
      if (!contains.path(variants, sp$vpath)) # Add to list, unless it already contains the same path
      {
        variants[[length(variants)+1]] <- list(g=newgraph, variants=list(path=sp$epath, vert=sp$vpath, dist=spd))
      }
    }
  }
  return(variants)
}

######################
# Main K Path Function
######################

k.shortest.paths <- function(graph, from, to, k, gmode){
  # First shortest path
  k0 <- get.shortest.paths(graph,from,to, output='both', mode=gmode)
  
  # Number of currently found shortest paths
  kk <- 1
  
  # List of alternatives
  variants <- list()
  
  # Shortest variants
  shortest.variants <- list(list(g=graph, path=k0$epath, vert=k0$vpath, dist=shortest.paths(graph,from,to,mode=gmode)))
  
  # Until k shortest paths are found
  while(kk<k){
    # Take last found shortest path
    last.variant <- shortest.variants[[length(shortest.variants)]]              
        
    # Calculate all alternatives
    variants <- calculate.variants(variants, last.variant, from, to, gmode)
        
    # Find shortest alternative
    sp <- select.shortest.path(variants)
    
    # Add to list, increase kk, remove shortest path from list of alternatives
    shortest.variants[[length(shortest.variants)+1]] <- list(g=variants[[sp]]$g, path=variants[[sp]]$variants$path, vert=variants[[sp]]$variants$vert, dist=variants[[sp]]$variants$dist)
    kk <- kk+1
    variants <- variants[-sp]
  }
  return(clean_output(shortest.variants))
}

```


#### Clean output from k shortest paths
```{r k_shortest, warning=FALSE}

clean_output <- function(k_shortest){
  
  shortest_routes <- setNames(data.frame(matrix(ncol = 4, nrow = 0)),
                              c("itinerary_id", "station", "line", "direction"))
  route <- vector(mode = "character")
  
  # loop through itineraries
  for(i in 1:k_num){
    station_list <- names(unlist(k_shortest[[i]]$`vert`))
    line_ids <- edge.attributes(mta_igraph, index =
                                unlist(k_shortest[[i]]$`path`))$route_ids
    prev_line_ids <- str_split(line_ids[1], "_")[[1]]
    dirs <- edge.attributes(mta_igraph, index =
                                unlist(k_shortest[[i]]$`path`))$direction_id
 
    # keep track of how many prev_line_ids to append
    r = 0
    
    # loop through stations
    for (j in 1:length(station_list)){
      
      shortest_routes <- shortest_routes %>% 
        add_row(itinerary_id = i, line =line_ids[j], 
                station = station_list[j], direction = dirs[j])
      
      # split line ids into a list
      cur_line_ids <- str_split(line_ids[j], "_")[[1]]
      
      # find the common lines between current and previous stops
      intersect_lines <- intersect(cur_line_ids, prev_line_ids)
      
      # check if you need to transfer
      if (length(intersect_lines) == 0 | "T" %in% cur_line_ids){
        
        if("T" %in% cur_line_ids){
          shortest_routes[j,]$direction <- shortest_routes[j-1,]$direction
        }
        
        else{
          # otherwise set to current line_ids
          intersect_lines <- cur_line_ids
        }
      
        # fill in all of the common lines up until the transfer
        route <- append(route, rep(paste(prev_line_ids, collapse = "_"), r))
        r = 0

      } # endif

      r = r+1
      prev_line_ids <- intersect_lines
      
    } # endfor
    
    route <- append(route, "END")
    
  } # endfor
  
  shortest_routes$line <- route
  
  return(shortest_routes)
}
```

#### Test k shortest paths function
```{r}
k_num <- 3

#Find k_num shortest paths leading going from stop R13 to stop D05 
k_shortest <- k.shortest.paths(mta_igraph, from = "R13", to = "D05", k=k_num, gmode="out")

k_shortest %>% View
```


#### Function to get formatted itineraries
```{r}
get_itinerary <- function(shortest_paths_df, num_ititeraries) {
  
  # new df for the formatted itineraries
  itinerary <- setNames(data.frame(matrix(ncol = 7, nrow = 0)),
                              c("itinerary_id", "station", "line", "direction", "leg", "event", "event_id"))
  
  # format each itinerary_id separately
  for (i in 1:num_ititeraries) {
    df <- shortest_paths_df %>%
      filter(itinerary_id == i)
    
    # current index
    k = 1
    
    # keep track of leg
    l = 1
    
    df <- df %>% mutate(leg = l, event = "travel")
    df$event[1] <- "start_trip"
    
    # compare current and next rows for whole itinerary
    while (k < nrow(df)){
      # next index
      j = k+1
      
      # identify transfers
      if (df$station[k] != df$station[j] & df$line[k] != df$line[j]) {
        
        # identify an 'implicit transfer (e.g. transfer 120->120 from 1 to 2/3)
        if (df$line[j] != "") {
          df <- df %>% add_row(itinerary_id = df$itinerary_id[j], station = df$station[j],
                line = df$line[k], direction = df$direction[k], leg = l, .after = k)
        }
        
        # identify an 'explicit' transfer (e.g. transfer R11->629 from N/R to 4)
        else {
          df$line[j] <- df$line[k] 
          df$leg[j] <- df$leg[k]
        }
        
        # make note of transfer events
        df$event[j] <- "start_transfer"
        df$event[j+1] <- "end_transfer"
        
        # start_transfer direction gets inherited from previous station
        df$direction[j] <- df$direction[k]
        
        # skip newly added row
        k = k+1
        
        # new leg starts with a transfer
        l = l+1

      }
      
      k = k+1
      df$leg[k] <- l
      
    }
    
    # very last row is redundant
    df <- df[1:nrow(df)-1,]
    df$event[nrow(df)] <- "end_trip" 
    df <- df %>% mutate(event_id = seq(1:nrow(df)))
  
    # put it all together
    itinerary <- rbind(itinerary, df)
  }
  
    return(itinerary)
}
```


#### Test itinerary function
```{r}
itin <- get_itinerary(k_shortest, k_num)
```
