---
title: "20180718_mta_triptime_distribution"
author: "Phoebe Nguyen, Akbar Mirza"
date: "July 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggthemes)

# load todd's subway_data
load(file = "../../data/todd_subway_realtime.RData")
# NOTE: How We Get Subway_data_2
# subway_data.2 <- subway_data %>%
#   group_by(realtime_trip_id) %>%
#   arrange(desc(departure_time)) %>%
#   mutate(arrival_time=lead(departure_time)) %>%
#   ungroup()
#
# save(subway_data.2, file = "../../data/todd_subway_realtime_2.RData")
load(file = "../../data/todd_subway_realtime_2.RData")



# load stop_times from GTFS Static
stop_times <- read_csv("../../data/google_transit_subway_static/stop_times.txt")

# load station_data from mta
station_data <- read_csv("http://web.mta.info/developers/data/nyct/subway/Stations.csv")

# get stop_id and stop_name fields, and create a stop_id with N and S appended
station_data <- station_data %>% group_by(`GTFS Stop ID`, `Stop Name`) %>%
  select(stop_id = `GTFS Stop ID`, stop_name = `Stop Name`) %>%
  mutate(stop_id_N = sprintf('%sN', stop_id), stop_id_S = sprintf('%sS', stop_id)) %>%
  gather(`stop_id_N`, `stop_id_S`, key = "stop_ids", value = "stop_id") %>%
  select(stop_name, stop_id)

################################################################################
# Getting Station Information
################################################################################
# Generate edges between stations
trips <- stop_times %>% extract(trip_id, "route", regex=".*_.*_(.*)\\.\\..*", remove = FALSE)
trip_edges <- trips %>% mutate(prev_stop_id = ifelse(lag(trip_id) == trip_id, lag(stop_id), NA))
edges <- trip_edges %>% select(route, stop_id, prev_stop_id) %>% distinct()
# create stations dataframe with line, stop_id, stop_name
stations <- edges %>% left_join(station_data, by = c("stop_id"="stop_id")) %>% select(line=route, stop_id, stop_name)
```

## Introducing One Stop Transfer

- Explore a slightly different route on this line that could use either the local or a combination of the local and express with a transfer in between(e.g., 18th to 96th st.)
- Compare the distribution of trip times using the local to those using the local + express

```{r compute-trip-time}

# NOTE: DEBUG CODE
start_station <- "120S"
start_line <- "2"
transfer_station <- "127S"
transfer_line <- "1"
end_station <- "131S"
start_time <- "2018-01-20 09:30:00"

compute_trip_time <- function (start_station, start_line, transfer_station, transfer_line, end_station, start_time) {

    # find first trips on the start_line at the start_station that depart after start_time
    first_trains <- subway_data.2 %>%
      filter(stop_mta_id == start_station | stop_mta_id == transfer_station,
             route_mta_id == start_line) %>%
      filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
      group_by(realtime_trip_id) %>%
      mutate(count = n()) %>%
      filter(count >= 2) %>%
      ungroup() %>%
      arrange(departure_time)
    # ungroup and arrange by departure_time

    # get the trip_id of the first trip after the depart_time
    first_trip_id <- first(first_trains$realtime_trip_id)

    # NOTE: arrival_time to first_train # 971896

    # find the arrival time for trip_id at transfer station
    transfer_arrival_time <- subway_data.2 %>%
      filter(realtime_trip_id == first_trip_id,
             stop_mta_id == transfer_station) %>%
      .$arrival_time

    # find the first trip on the transfer_line at the transfer_station that departs after the arrival_time of our express train
    transfer_trains <- subway_data.2 %>%
      filter(stop_mta_id == transfer_station | stop_mta_id == end_station,
             route_mta_id == transfer_line) %>%
      filter(ymd_hms(departure_time) > ymd_hms(transfer_arrival_time)) %>%
      group_by(realtime_trip_id) %>%
      mutate(count = n()) %>%
      filter(count >= 2) %>%
      ungroup() %>%
      arrange(departure_time)
      # note above

    transfer_trip_id <- first(transfer_trains$realtime_trip_id)

    # NOTE: arrival_time to transfer_train

    # compute the time difference of when our train arrives at end_station and start_time
    dest_arrival_time <- subway_data.2 %>%
      filter(realtime_trip_id == transfer_trip_id,
             stop_mta_id == end_station) %>%
      .$arrival_time

    time_diff <- as.numeric(difftime(ymd_hms(dest_arrival_time), ymd_hms(start_time)), units = "mins")

     return(time_diff)
#
}
#
compute_trip_time("120S", "2", "127S", "1", "131S", "2018-01-22 09:30:00")

```

## Comparing Distributions of Local vs Local + Express

### Computing Local + Express Wait Times On Large Scale Efficiently

```{r example-for-trips}

compute_itenerary_pair <- function(start_lines, end_lines, start_id, transfer_start_id, transfer_end_id, end_id, start_time) {

  # TODO: filter subway_data to only trips within a certain time_interval

  # get train information for start_lines at start_id
  leg1_data <- subway_data.2 %>%
    filter(stop_mta_id == start_id | stop_mta_id == transfer_start_id,
           # TODO: modify filter to take multiple lines
           route_mta_id == start_lines) %>%
    group_by(realtime_trip_id) %>%
    mutate(trip_start_time = min(departure_time)) %>%
    # get rid of records for extra record at start stops
    filter(n() >= 2, stop_mta_id == transfer_start_id) %>%
    arrange(arrival_time)

  leg2_data <- subway_data.2 %>%
    filter(stop_mta_id == transfer_end_id | stop_mta_id == end_id,
            # TODO: modify filter to take multiple lines
            route_mta_id == end_lines) %>%
    group_by(realtime_trip_id) %>%
    mutate(trip_end_time = max(departure_time)) %>%
    # get rid of records for extra record at start stops
    filter(n() >= 2, stop_mta_id == transfer_end_id) %>%
    arrange(departure_time)


  leg1_ptr <- 1
  leg2_ptr <- 1
  # create vectors to store new info
  diffs <- vector('numeric')
  start_times <- vector('double')
  transfer_arrival_times <- vector('double')
  end_times <- vector('double')
  transfer_departure_times <- vector('double')
  realtime_start_ids <- vector('numeric')
  realtime_end_ids <- vector('numeric')

  while (leg1_ptr < nrow(leg1_data)) {

    if (leg2_data$departure_time[leg2_ptr] > leg1_data$arrival_time[leg1_ptr]) {
      # populate fields that will be used for our returned data_frame
      realtime_start_ids <- c(realtime_start_ids, leg1_data$realtime_trip_id[leg1_ptr])
      start_times <- c(start_times, leg1_data$trip_start_time[leg1_ptr])
      transfer_arrival_times <- c(transfer_arrival_times, leg1_data$arrival_time[leg1_ptr])
      realtime_end_ids <- c(realtime_end_ids, leg2_data$realtime_trip_id[leg2_ptr])
      transfer_departure_times <- c(transfer_departure_times, leg2_data$departure_time[leg2_ptr])
      end_times <- c(end_times, leg2_data$trip_end_time[leg2_ptr])
      diffs <- c(diffs,
                 as.numeric(difftime(leg2_data$trip_end_time[leg2_ptr],
                                     leg1_data$trip_start_time[leg1_ptr]),
                            units = "mins"))
      
      leg1_ptr = leg1_ptr + 1
    } else {
      leg2_ptr = leg2_ptr + 1
    }
  
  }

  new_df <- data.frame(start_trip_id = realtime_start_ids,
                       starts = as.POSIXct(start_times, origin = "1970-01-01", tz = "UTC"),
                       transfer_starts = as.POSIXct(transfer_arrival_times, origin = "1970-01-01", tz = "UTC"),
                       end_trip_id = realtime_end_ids,
                       transfer_ends = as.POSIXct(transfer_departure_times, origin = "1970-01-01", tz = "UTC"),
                       ends = as.POSIXct(end_times, origin = "1970-01-01", tz = "UTC"),
                       diffs = diffs)

  return(new_df)
}

local_line <- "1"
express_line <- "2"
start_stop_id <- "120S"
transfer_stop_id <- "127S"
end_stop_id <- "131S"
start_time <- "2018-01-20 09:30:00"

# Getting 2 Trains
data_2_train <- subway_data.2 %>%
  filter(stop_mta_id == "120S" | stop_mta_id == "127S",
         route_mta_id == "2") %>%
  # filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
  group_by(realtime_trip_id) %>%
  mutate(start_time = min(departure_time)) %>%
  filter(n() >= 2, stop_mta_id == "127S") %>%
  arrange(arrival_time)

# Getting 1 Trains
data_1_train <- subway_data.2 %>%
  filter(stop_mta_id == "127S" | stop_mta_id == end_stop_id,
         route_mta_id == local_line) %>%
  # filter(ymd_hms(departure_time) > ymd_hms(start_time)) %>%
  group_by(realtime_trip_id) %>%
  mutate(end_time = max(departure_time)) %>%
  filter(n() >= 2, stop_mta_id == "127S") %>%
  arrange(departure_time)


# NOTE: train 1 start time and train 2 end time for debugging and evaluating purposes
i = 1
j = 1
times <- vector('numeric')
start_times <- vector('double')
transfer_arrival_times <- vector('double')
end_times <- vector('double')
transfer_departure_times <- vector('double')
realtime_start_ids <- vector('numeric')
realtime_end_ids <- vector('numeric')

while (i < nrow(data_2_train)) {
  if (data_1_train$departure_time[j] > data_2_train$arrival_time[i]) {
    # df <- data.frame(trip_time = 0, start_trip = 0, end_trip = 0)
    realtime_start_ids <- c(realtime_start_ids, data_2_train$realtime_trip_id[i])
    start_times <- c(start_times, data_2_train$start_time[i])
    transfer_arrival_times <- c(transfer_arrival_times, data_2_train$arrival_time[i])
    realtime_end_ids <- c(realtime_end_ids, data_1_train$realtime_trip_id[j])
    transfer_departure_times <- c(transfer_departure_times, data_1_train$departure_time[j])
    end_times <- c(end_times, data_1_train$end_time[j])
    times <- c(times, as.numeric(difftime(data_1_train$end_time[j], data_2_train$start_time[i]), units = "mins"))
    i = i + 1
  } else {
    j = j + 1
  }
}

# verify with real data
new_df <- data.frame(start_trip_id = realtime_start_ids,
                     starts = as.POSIXct(start_times, origin = "1970-01-01", tz = "UTC"),
                     transfer_starts = as.POSIXct(transfer_arrival_times, origin = "1970-01-01", tz = "UTC"),
                     end_trip_id = realtime_end_ids,
                     transfer_ends = as.POSIXct(transfer_departure_times, origin = "1970-01-01", tz = "UTC"),
                     ends = as.POSIXct(end_times, origin = "1970-01-01", tz = "UTC"),
                     diffs = times)

# Views to Verify Trips
# ORIGINAL SUBWAY DATA
subway_data %>% filter(realtime_trip_id == 3300 & stop_mta_id == "120S" | realtime_trip_id == 3327 & stop_mta_id == "131S") %>% View("subway_data 3300")
# NEW DATA FRAME DATA
new_df %>% filter(start_trip_id == 3300) %>% View("new_df 3300")

test_df <- compute_itenerary_pair("2", "1", "120S", "127S", "127S", "131S", start_time)
# TODO: figure out multiple transfers with multiple lists

# PSEUDOCODE
# # NOTE: adding third train to the loop
# # parsing two rows for each train 
# i = 1
# j = 1
# k = 1
# counter_list <- c(i,j,k)


test_df <- data.frame(d=1)

trains <- c("1", "2", "L")

stop_station <- list(c("120S", "127S"),
                      c("127S", "132S"),
                      c("L02S", "L06S"))

counters <- rep(1, length(trains))

# create data frame that we're going to use
# test <- data.frame("itin_id" = c("1","1","1"), 
#                    "station"= c("120S", "127S","127S", "132S", "L02S", "L06S"),
#                    "line" = c("1", "1","2","2","L","L"))

# add column_names that are needed
for (train in trains) {
  test_df[[paste(train, 'train_id', sep='_')]] <- NA
  test_df[[paste(train, 'start_time', sep='_')]] <- NA
  test_df[[paste(train, 'end_time', sep='_')]] <- NA
}

test_df <- subset(test_df, select = -c(d)) # %>% na.omit
# generate the values needed for each as a vector and rbind to add "rows" to our df

# list of data frames
len <- length(trains)
train_data <- list()

for (i in 1:len) {
  
  start <- stop_station[[i]][1]
  end <- stop_station[[i]][2]
  
  # TODO: filter subway_data to get appropriate information for each train
  leg_data <- subway_data.2 %>%
    filter(stop_mta_id == start | stop_mta_id == end,
           # TODO: modify filter to take multiple lines
           route_mta_id == trains[i]) %>%
    group_by(realtime_trip_id) %>%
    mutate(start_time = min(departure_time)) %>%
    # get rid of records for extra record at start stops
    filter(n() >= 2, stop_mta_id == end) %>%
    arrange(arrival_time)
  # append to list
  print(nrow(leg_data))
  train_data[[i]] <- leg_data
}

diffs <- vector('numeric')
start_times <- vector('double')
transfer_arrival_times <- vector('double')
end_times <- vector('double')
transfer_departure_times <- vector('double')
realtime_start_ids <- vector('numeric')
realtime_end_ids <- vector('numeric')

i <- 1

row_info <- list()

while (i < nrow(train_data[[1]])) {
  train_num <- trains[1]
  # df[sprintf("%s_train_id"), train_num] <- train_data[1]$realtime_trip_id[i]
  # df[sprintf("%s_start_time"), train_num] <- train_data[1]$start_time[i]
  end_time <- train_data[[1]]$departure_time[i]
  # df[sprintf("%s_end_time"), train_num] <- end_time
  row_info <- c(row_info, train_data[[1]]$realtime_trip_id[i], train_data[[1]]$start_time[i], train_data[[1]]$departure_time[i])
  
  for (ii in 2:length(trains)) {
    while (train_data[[ii]]$departure_time[counters[ii]] < end_time) {
      counters[ii] = counters[ii] + 1
    }
    train_num_ii <- trains[ii]
    curr_idx <- counters[ii]
    # df[sprintf("%s_train_id"), train_num_ii] <- train_data[ii]$realtime_trip_id[curr_idx]
    # df[sprintf("%s_start_time"), train_num_ii] <- train_data[ii]$start_time[curr_idx]
    # end_time_ii <- train_data[ii]$departure_time[curr_idx]
    # df[sprintf("%s_end_time"), train_num_ii] <- end_time_ii
    row_info <- c(row_info, train_data[[ii]]$realtime_trip_id[curr_idx], train_data[[ii]]$start_time[curr_idx], train_data[[ii]]$departure_time[curr_idx])
  }
  test_df <- rbind(test_df, row_info)
  row_info <- list()
  # increment i
  i = i + 1
}

# TODO: FIGURE OUT WHY THERE ARE DUPLICATES
  
  #  for ii in range(2, len(counter_list)+1):
#    while data[ii][counter_list[ii]] < arrival_time:
#      counter_list[ii] += 1
#    #record when train ii leaves
#    arrival_time = data[ii][counter_list[ii]]$end_time
#    #record when train ii departs


# #list of start_times 
# for (x in counter_list) {
#  add start_time_column to list of start times
#  add end_time_column to list of end times
#  add train_id column to list of train ids
# }
# 
# times <- vector('numeric')
# start_times <- vector('double')
# # transfer_arrival_times.1 <- vector('double')
# # transfer_arrival_times.2 <- vector('double')
# end_times <- vector('double')
# final_df <- c(start_time_train1, start_time_train2, start_time_train3)
# # transfer_departure_times <- vector('double')
# realtime_start_ids <- vector('numeric')
# realtime_end_ids <- vector('numeric')
# 
# data is a list of dataframes, where data[i] corresponds to the ith train in the itinerary
# 
# while (i < nrow(data_1_train)) {
#  record when train i leaves
#  arrival_time = data_1[i]$end_time #record when train i arrives
#  #put this in arrival_time_list for the 1 train
#  for ii in range(2, len(counter_list)+1):
#    while data[ii][counter_list[ii]] < arrival_time:
#      counter_list[ii] += 1
#    #record when train ii leaves
#    arrival_time = data[ii][counter_list[ii]]$end_time
#    #record when train ii departs
# }

```