---
title: "Map With Transfers and Igraph"
author: "Graph Team"
date: "July 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Sasha, Ayliana, Brian, & Taxi*  

#### Load libraries
```{r libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggmap)
library(leaflet)
library(igraph)
library(scales)
```


#### Load MTA data & Google Transit static files
```{r load_data, warning=FALSE, message=FALSE}
stations <- read_csv('http://web.mta.info/developers/data/nyct/subway/Stations.csv')
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
trips <- read_csv('../../data/google_transit_subway_static/trips.txt')
stop_times <- read_csv('../../data/google_transit_subway_static/stop_times.txt')
transfers <- read_csv('../../data/google_transit_subway_static/transfers.txt')

load('../../data/todd_subway_realtime.RData')
```


#### The `stations` data looks like:
```{r stations}
as.data.frame(stations) %>% head
```


#### We might want to view each route on its own line
```{r warning=FALSE}
new_stations <- stations %>% 
  select(complex_id = `Complex ID`, stop_id = `GTFS Stop ID`, stop_name = `Stop Name`, 
         stop_lat = `GTFS Latitude`, stop_lon = `GTFS Longitude`, line=Line, daytime_routes=`Daytime Routes`)

new_stations <- new_stations %>% 
  separate(daytime_routes, into=c('route1', 'route2', 'route3', 'route4'), sep=" ") 

ss1 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route1)
ss2 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route2)
ss3 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route3)
ss4 <- new_stations %>% select(complex_id, stop_id, stop_name, stop_lat, stop_lon, line, route=route4)

routes_by_lines <- rbind(ss1, ss2, ss3, ss4) %>% filter(!is.na(route))

as.data.frame(routes_by_lines) %>% head
```


#### Use `stop_times` to understand the schedules/sequences of each line
```{r}
connections <- stop_times %>%
  left_join(stops) %>% 
  extract(trip_id, c("route_id"), regex=".*_.*_([^.]*)\\.\\.?.*", remove=FALSE) %>%
  mutate(stop_id = substr(stop_id, 1, 3),
    prev_stop_id = ifelse(trip_id == lag(trip_id), lag(stop_id), NA),
    prev_stop_lat = ifelse(trip_id == lag(trip_id), lag(stop_lat), NA),
    prev_stop_lon = ifelse(trip_id == lag(trip_id), lag(stop_lon), NA),
    prev_stop_name = ifelse(trip_id == lag(trip_id), lag(stop_name), NA))

sequences <- connections %>%
  left_join(routes) 

# catch NA color values
sequences$route_color <- replace_na(sequences$route_color, "6D6E71") 

sequences <- sequences %>%
  mutate(route_color = sprintf("#%s", route_color)) %>%
  select(route_id, stop_id, stop_name, stop_lat, stop_lon, 
         prev_stop_id, prev_stop_name, prev_stop_lat, prev_stop_lon,
         route_color) %>%
  distinct 

```


#### Use `transfers` to understand where we can transfer to other lines
```{r transfers}

#Prepping the transfer data
transfer_sequences <- transfers %>% left_join(stops, by = c("to_stop_id" = "stop_id")) %>% 
  left_join(stops, by = c("from_stop_id" = "stop_id")) %>%
  mutate(route_id = "T", route_color = "#000000") %>%
  select(route_id, stop_id = to_stop_id, stop_name = stop_name.x, stop_lat = stop_lat.x, stop_lon = stop_lon.x, prev_stop_id = from_stop_id, prev_stop_name = stop_name.y, prev_stop_lat = stop_lat.y, prev_stop_lon = stop_lon.y, route_color, weight = min_transfer_time)

```


#### Get weights for the connections between stations
```{r weights, warning=FALSE}

station_weights <- realtime %>%
  mutate(stop_mta_id = substr(stop_mta_id, 1, 3)) %>%
  arrange(realtime_trip_id, departure_time) %>% 
  mutate(travel_time = ifelse(realtime_trip_id == lag(realtime_trip_id), departure_time - lag(departure_time), NA)) %>%
  mutate(prev_stop_mta_id = ifelse(realtime_trip_id == lag(realtime_trip_id), lag(stop_mta_id), NA)) %>%
  filter(!is.na(travel_time)) %>% 
  group_by(route_mta_id, stop_mta_id, prev_stop_mta_id) %>% 
  summarize(weight = mean(travel_time))

full_sequences <- sequences %>% left_join(station_weights, by = c("stop_id" = "stop_mta_id", "prev_stop_id" = "prev_stop_mta_id", "route_id" = "route_mta_id"))

#Transfers already has weights from the min_transfer_time field (now renamed weight) so we rbind the two sequence objects

full_sequences <- rbind(full_sequences, transfer_sequences)

```

#### Make a subway map!
```{r ggmap, warning=FALSE}

nyc_map <- get_map(location = c(lon = -73.9568247, lat = 40.7202688), maptype = "terrain", zoom = 12)
ggmap(nyc_map) +
 geom_point(data = full_sequences, aes(x = stop_lon, y = stop_lat)) +
 geom_segment(data = full_sequences, aes(x=prev_stop_lon, y=prev_stop_lat, xend=stop_lon, yend=stop_lat, color = route_id))

```


#### Make another (prettier) map
```{r leaflet}

map <- leaflet() %>%
  addTiles() %>%
  setView(-73.95, 40.72, zoom = 10.5) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(sequences$stop_lon, sequences$stop_lat, color = sequences$route_color,
                   radius = 4, stroke = FALSE, fillOpacity = 0.7)

for (i in 1:nrow(full_sequences)) {
  map <- map %>%
    addPolylines(lat = c(full_sequences[i,]$stop_lat, full_sequences[i,]$prev_stop_lat),
                 lng = c(full_sequences[i,]$stop_lon, full_sequences[i,]$prev_stop_lon),
                 color = full_sequences[i,]$route_color,
                 dashArray = ifelse(full_sequences[i,]$route_id == "T", c(8), c(0)),
                 weight = 1)
}

map
```


#### Map with walking grid and stations (for presentation)
```{r}
library(mapview)

# sequences$stop_lat <- jitter(sequences$stop_lat, factor = 10)
# sequences$stop_lon <- jitter(sequences$stop_lon, factor = 10)

map <- leaflet() %>%
  addTiles() %>%
  setView(-73.95, 40.72, zoom = 10.5) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(data = filtered_grid, lng = ~lon, lat = ~lat,
                     color = "#000000",
                     radius = 0.5, stroke = FALSE, fillOpacity = 0.3) %>%
  addCircleMarkers(sequences$stop_lon, sequences$stop_lat, color = sequences$route_color,
                   radius = 4, stroke = FALSE, fillOpacity = 0.7)

mapshot(map, url = "stations_and_walking_grid.html", file = "stations_and_walking_grid.png")

```

#### Make an igraph
```{r igraph}

theme_set(theme_bw())

igraph_edges <- full_sequences %>% filter(!is.na(prev_stop_id), !is.na(weight)) %>% select(prev_stop_id, stop_id, weight)
mta_igraph <- graph.data.frame(igraph_edges, directed=TRUE)

plot(mta_igraph) 

```

#### Find shortest route
```{r shortest_function}

shortest_path <- shortest_paths(mta_igraph, from = c("R19"), to = c("104"), mode="in")

shortest_route <- data.frame(stop_id = (sapply(shortest_path$vpath, as_ids)), stringsAsFactors = FALSE) %>% left_join(stops, by = "stop_id") %>% select(stop_id, stop_name)

shortest_route

```

