---
title: "20180727_mta_triptime_distribution"
author: "Phoebe Nguyen, Akbar Mirza"
date: "July 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)
library(ggthemes)
library(ROCR)
library(here)
```
```{r}
load(file = here("data", "todd_subway_realtime.RData"))
source(file = here("src", "itinerary_evaluation.R"))
```

## Plotting Functions

```{r function-definitions}
compute_auc <- function(itin_pair_df) {
  # pull out the number from the ROCR package with pred <- predictions(predicted, actual); performance(pred, 'auc')@y.values[[1]]
  # or  with the pROC package: as.numeric(auc(response = actual, predictor = predicted))
  # itinerary is the actual value, trip time is the predicted value
  auc_df <- itin_pair_df %>%
    mutate(itin_line = as.factor(paste0(group, itin_id, sep=''))) %>%
    ungroup() %>%
    select(itin_line, time_diff)
  
  pred <- prediction(as.numeric(auc_df$time_diff, units = "mins"), as.factor(auc_df$itin_line))
  perf_nb <- performance(pred, 'auc')@y.values[[1]]
  perf <- performance(pred, 'auc')
  ifelse( perf_nb < 0.5, return(1 - perf_nb), return(perf))
}

# given a df with historical time_diffs for many initineraries
# will return a plot that maps avg time_diff vs variance
plot_frontier <- function(itins_df) {

  plot_data <- itins_df %>%
    #mutate(day_of_week = weekdays(leg1_start_time), hour = hour(leg1_start_time)) %>%
    # filter out weekends, and only get hour_start, hour_end
    #filter(isWeekend(day_of_week) == F, hour >= 7 & hour < 10) %>%
    group_by(itin_id, label) %>%
    mutate(avg = mean(time_diff), stdv = sd(time_diff))

  plot <- plot_data %>% ggplot(aes(x = stdv, y = avg)) +
    geom_point() +
    xlab('Standard Deviation of Trip Time') +
    ylab('Average Trip Time') +
    geom_text(aes(label = paste0(group, itin_id,sep='')), hjust = 0.5, vjust = 1)

  return(plot)
}
```

## Evaluating Interesting Multi-Leg Trips

Evaluating Trip Itineraries With Multiple Legs.

### Red Line vs Orange Line

Looking at trip time distributions between 96 St and 23 St on the Red Line (1, 2, 3) versus the Orange Line (B, D) 

#### Gather Red Line Data

```{r red-line-data}
red_line_data <- read_csv(here::here("data", "itineraries-Red-Line-96-23-train.csv"))

# filter out event and adding direction to station id
red_line_cleaned <- red_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
red_line_itins <- compute_all_itins(red_line_cleaned, subway_data)%>%
  mutate(group='red')
```

#### Gather Orange Line Data

```{r orange-line-data}
orange_line_data <- read_csv(here::here("data", "itineraries-Orange-Line-96-23-train.csv"))

# filter out event and adding direction to station id
orange_line_cleaned <- orange_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
orange_line_itins <- compute_all_itins(orange_line_cleaned, subway_data) %>%
  mutate(group='orange')

#combine first line df with second line df
red_orange_df <- bind_rows(red_line_itins, orange_line_itins)
```

#### Evaluating Differences Between Red & Orange Line & Plotting

```{r red-orange-comparison}
top_2_trips <- red_orange_df %>% group_by(group) %>% filter(itin_id == "1")
  
# Plot Distribution as Densities
plot_densities(top_2_trips)

# Plot Frontiers
plot_frontier(red_orange_df)

# Plot AUC for Top 2 Trips
compute_auc(top_2_trips)
```

### Midtown to Jamaica (JFK) during afternoon rush hour (3pm-5pm)

```{r midtown-jamaica-plots}

# load midtown-jamaica data, and clean for our uses
#   - filter out event and add direction to station id
mj_data <- read_csv(here::here("data", "itineraries-Midtown-Jamaica.csv")) %>%
  filter(event != "travel", itinerary_id < 2)

mj_df <- compute_all_itins(mj_data, subway_data, 15, 17) %>% 
  filter(time_diff != 0) 

# filtering out 0's from E train
plot_densities(mj_df)

# NOTE: DEBUG CODE
# subway_data %>% filter(stop_mta_id == "F12N" | stop_mta_id == "G06N", route_mta_id == "E") %>% group_by(realtime_trip_id) %>%
#   mutate(start_time = min(departure_time), count = n()) %>%
#   ungroup() %>% filter(realtime_trip_id %in% c(13391, 13404, 13457, 13495, 13517)) %>% View
```
### Midtown to Mets Willets (LGA) during afternoon rush hour (3pm-5pm)

```{r Bryant Park- Mets Willets}
bm_data <- read_csv(here::here("data","itineraries-5ave-MetsWillets.csv")) %>%
  filter(event != "travel", itinerary_id < 2)

bm_df <- compute_all_itins(bm_data, subway_data, 15,17) %>% filter(time_diff > 2) # filtering out 0's from E train
plot_densities(bm_df)


```

### Red/Green to Fulton

#### Gathering Red Line Data from 96 - Fulton
```{r red-96-fulton-data}
red_line_data <- read_csv(here::here("data", "itineraries-96-Fulton-red.csv"))

# filter out event and adding direction to station id
red_line_cleaned <- red_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
red_line_itins <- compute_all_itins(red_line_cleaned, subway_data) %>%
  mutate(group='red')
```

#### Gathering Green Line Data from 96 - Fulton
```{r green-96-fulton-data}
green_line_data <- read_csv(here::here("data", "itineraries-96-Fulton-green.csv"))

# filter out event and adding direction to station id
green_line_cleaned <- green_line_data %>%
  filter(event != "travel", itinerary_id < 4)

#output for first data input
green_line_itins <- compute_all_itins(green_line_cleaned, subway_data) %>%
  mutate(group='green')
```

#### Comparison of Red v. Green Data from 96 - Fulton
```{r red-green-96-fulton-comparison}
#combine first line df with second line df
red_green_df <- bind_rows(red_line_itins, green_line_itins)

red_green_top_pair <- red_green_df %>% group_by(group) %>% filter(itin_id == "1")
  
# Plot Distribution as Densities
plot_densities(red_green_top_pair)

# Plot Frontiers
plot_frontier(red_green_df)

# Plot AUC for Top 2 Trips
compute_auc(red_green_top_pair)
```

## Plots for Debugging

```{r debugging_plots}

# TODO: Write A Function To Show Trains On Plot

trains = c("1", "2_3")
start_stop = c("120S", "123S")
end_stop = c("123S", "229S")
train_data <- vector("list", length(trains))

for (i in 1:length(trains)) {
  start <- start_stop[i]
  end <- end_stop[i]
  lines <- trains[i]
  
  leg_data <- get_leg_data(lines, start, end, subway_data)
  
  # NOTE: DEBUG CODE
  # print(nrow(leg_data))
  
  # append to list
  train_data[[i]] <- leg_data
}

# plot_data <- bind_rows(train_data)
# plot_data %>% View

debug_train_schedules_plot <- function(train_data, subway_data, start_hour = 8, end_hour = 9) {
  # filtered_df <- itins_df %>% filter(hour(leg1_start_time) >= start_hour & hour(leg1_start_time) <= end_hour)
  
  plot_data <- rbind_list(train_data)
  print(plot_data)
  plot_data %>%
    filter(month(start_time) == 2,
           day(start_time) == 16,
           hour(start_time) >= start_hour & hour(start_time) <= end_hour) %>%
    arrange(start_time) %>%
    mutate(itin = row_number()) %>%
    ggplot(aes(color = route_mta_id)) +
    geom_segment(aes(x = start_time, xend = departure_time, y = itin, yend = itin)) +
    geom_text(aes(label = realtime_trip_id, x = start_time, y = itin))
  
}

debug_train_schedules_plot(train_data, subway_data)

# NOTE: DEBUG PLOT - show all trains individually on plot

debug_itin_plot <- function(itin_df, subway_data, itin_number, num_trips = 5, start_hour = 8, end_hour = 9) {
  
  filtered <- itin_df %>% filter(itin_id == itin_number)
  
  reshaped_df <- filtered %>% select(matches("leg*")) %>% ungroup() %>%
    mutate(row_id = 1:n()) %>%
    gather(variable, value, -row_id) %>%
    extract(variable, c("leg", "var"), 'leg(.*?)_(.*)') %>%
    select(row_id, leg, var, value) %>%
    spread(var, value) %>%
    na.omit() %>%
    filter(row_id <= num_trips)
  
  reshaped_df$start_time <- as.POSIXct(as.numeric(reshaped_df$start_time), origin = "1970-01-01", tz = "UTC")
  reshaped_df$end_time <- as.POSIXct(as.numeric(reshaped_df$end_time), origin = "1970-01-01", tz = "UTC")
    
  
  plot <- reshaped_df %>% ggplot(aes(x = start_time, y = row_id)) +
    geom_segment(aes(xend = end_time, yend = row_id, color = as.factor(line))) +
    geom_text(aes(label = trip_id)) 
  
  return(plot)
}

debug_itin_plot(green_line_itins, subway_data, 1)

# filtered_test_df <- test_df %>% filter(hour(leg1_start_time) >= 8 & hour(leg1_start_time) <= 9)

# plot_data <- bind_rows(train_data[[1]], train_data[[2]], train_data[[3]])
# plot_data %>% filter(month(start_time) == 2 & day(start_time) == 16 & hour(start_time) > 19 & hour(start_time) < 24) %>%
#   arrange(start_time) %>%
#   mutate(itin = row_number()) %>%
#   ggplot(aes(color = route_mta_id)) +
#   geom_segment(aes(x = start_time, xend = departure_time, y = itin, yend = itin)) +
#   geom_text(aes(label = realtime_trip_id, x = start_time, y = itin))

# TODO: Write A Function To Show Itineraries On Plot

# ITINERARY PLOT
 red_green_top_pair %>% mutate(itin = row_number()) %>%
   filter(itin <= 5) %>%
   ggplot() +
     geom_segment(aes(x = leg1_start_time, xend = leg1_end_time, y = itin, yend = itin), color = 'red') +
     geom_segment(aes(x = leg2_start_time, xend = leg2_end_time, y = itin, yend = itin), color = 'blue') +
     geom_segment(aes(x = leg3_start_time, xend = leg3_end_time, y = itin, yend = itin), color = 'green') +
     geom_text(aes(label = leg1_trip_id, x = leg1_start_time, y = itin), color = "red") +
     geom_text(aes(label = leg2_trip_id, x = leg2_start_time, y = itin), color = "blue") +
     geom_text(aes(label = leg3_trip_id, x = leg3_start_time, y = itin), color = "green") +
     ylim(c(0, 10))

# filtered_test_df %>% ggplot() +
#   geom_histogram(aes(x = time_diff))
```