---
title: "L train 2"
author: "Sasha, Ayliana, Taxi"
date: "July 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Load libraries
```{r libs, warning=FALSE, message=FALSE}
library(igraph)
library(leaflet)
library(RColorBrewer)
library(geosphere)
library(sp)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(ggmap)
library(KernSmooth)
library(ggplot2)
library(gstat)
library(raster)
library(mapview)
library(tidyverse)
```

#### Load data
```{r load-data, message= FALSE}
stops <- read_csv('../../data/google_transit_subway_static/stops.txt')
stops <- stops %>% select(stop_id, stop_lat, stop_lon, stop_name)

routes <- read_csv('../../data/google_transit_subway_static/routes.txt')
routes$route_color <- replace_na(routes$route_color, "000000") 
routes <- routes %>%
  mutate(route_color = sprintf("#%s", route_color))

source('../../src/path_finding.R')
load('../../data/igraph_edges.rdata')

r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)
```


#### distance function for summing graph weights
```{r dist1}
distance <- function(graph, path) sum(E(graph, path=path)$weight)
```



#### distance function for walking to stations (in km)
```{r dist2}
Distance <- function(LatA, LonA, LatB, LonB){
  scaling_factor <- 110.25
  x <- LatA - LatB
  y <- (LonA - LonB) * cos(40.75 * (pi)/180)

  return (scaling_factor*sqrt(x * x + y * y))

}
```


#### Get map coords (once we have the itinerary)
```{r get-map-coords}
# retrieve coordinates and stop names for itineraries
get_map_data <- function(itinerary) {
  
  map_data <- itinerary %>%
  left_join(stops, by = c("station" = "stop_id", "stop_name" = "stop_name"))  %>%
  group_by(itinerary_id) %>%
  mutate(prev_station = lead(station), prev_stop_name = lead(stop_name),
         prev_stop_lat = lead(stop_lat), prev_stop_lon = lead(stop_lon)) %>%
  extract(line, c("route_id"), regex="([^_][^_]?)_?.*", remove=FALSE) %>%
  ungroup() %>%
  left_join(routes) %>%
  select(itinerary_id, station, line, weight, leg, event, event_id,
         stop_name, stop_lat, stop_lon, prev_station, prev_stop_name, prev_stop_lat,
         prev_stop_lon, route_color)
  
  return(map_data)
}
```


#### layered map function
```{r layered-map}
map_an_itinerary <- function(home, start, itinerary) {
  
  black <- "#000000"
  stop_rad <- 5
  line_weight <- 3
  
  map_df <- get_map_data(itinerary) %>%
    mutate(line = str_replace_all(line, "_", " "),
           mins = as.integer(weight)%/%60)

  # map_df %>% View
  
  num_layers <- max(map_df$itinerary_id)

  map <- leaflet() %>%
    addTiles() %>%
    setView(median(map_df$stop_lon), median(map_df$stop_lat), zoom = 12) %>%
    addProviderTiles("CartoDB.Positron")

for (i in 1:num_layers) {
  df <- map_df %>% filter(itinerary_id==i)

  map <- map %>%
    # add station markers
    addCircleMarkers(df$stop_lon, df$stop_lat, color = df$route_color,
                     popup = paste("<b>", df$stop_name, "</b>", "<br/>", df$line),
                     radius = stop_rad, stroke = FALSE, fillOpacity = 0.7,
                     group= as.character(i)) %>%
    # add start marker
    addCircleMarkers(lat = home$lat, lng = home$lon, color = green, radius = 8, opacity = 0.9,
                     popup = paste("<b>", "start location", "</b>", "<br/>",
                                   home$lat, home$lon)) %>%
    # add walk to first station
    addPolylines(lat = c(home$lat, df$stop_lat[nrow(df)]),
                 lng = c(home$lon, df$stop_lon[nrow(df)]),
                 dashArray = c(8),
                 color = black, weight = line_weight,
                 popup = paste("<b>", "walk", "</b>", "<br/>", start$walking_mins, "mins")) %>%
    # add end marker
    addCircleMarkers(lat = df$stop_lat[1], lng = df$stop_lon[1], color = red, radius = 8, opacity = 0.9,
                     popup = paste("<b>", "end location", "</b>", "<br/>",
                                   df$stop_name[1]))

  for (j in 1:nrow(df)) {
    if (j == 1 & df$event[j] == "transfer") {
      map <- map %>% 
        addPolylines(lat = c(df$stop_lat[j], df$prev_stop_lat[j]),
                     lng = c(df$stop_lon[j], df$prev_stop_lon[j]),
                     dashArray = c(8), 
                     color = black, 
                     weight = line_weight,
                     popup = paste("<b>", "walk", "</b>", "<br/>", df$mins[j], "mins"))
    } else {
    map <- map %>%
      addPolylines(lat = c(df$stop_lat[j], df$prev_stop_lat[j]),
                   lng = c(df$stop_lon[j], df$prev_stop_lon[j]),
                   color = ifelse(df$event[j] == "start_transfer", black, df$route_color[j]),
                   popup = ifelse(df$event[j] == "start_transfer",
                                  paste("<b>", "transfer", "</b>",
                                        df$line[j+1], "to", df$line[j], "<br/>",
                                        df$mins, "mins"),
                                  paste("<b>", df$line[j], "</b>", "<br/>", df$mins[j], "mins")),
                   dashArray = ifelse(df$event[j] == "start_transfer", c(8), c(0)),
                   weight = line_weight,
                   group = as.character(i))
    
    }
  }

}

  map <- map %>%
    addLayersControl(overlayGroups = as.character(seq(1:num_layers)),
                     options = layersControlOptions(collapsed = FALSE))
  return(map)
}
```


### GET GRAPHS
#### with L
```{r}
igraph_edges <- mutate(igraph_edges, 'weight'=`90%`)
graph <- graph.data.frame(igraph_edges)
```

#### without L
```{r}
no_l_edges <- igraph_edges %>%
  filter(route_id != "L")
l_graph <-graph.data.frame(no_l_edges)
```


#### settings
```{r}
# final destination station
dest <- "A31N" # 14th St ******
# dest <- "103N"

# 5 mins walking from 96 St
home_lat <- 40.79613
home_lon <- -73.97688

# near L/canarsie
home_lat <- 40.651501
home_lon <- -73.895612

# metropolitan ave
home_lat <- 40.712037 
home_lon <- -73.894585 

# starting location ("home")
home <- data.frame(lat = c(home_lat), lon = c(home_lon))

# number of paths
k <- 1

# start/stop marker colors
green <- "#468504"
red <- "#cf3400"
```


```{r}
get_map_with_walking <- function(home, igraph) {
  map_data <- get.all.shortest.paths(igraph, dest,  mode = "out")

  map_data <- map_data$res %>% 
    lapply(function(x)data.frame(stop_id = names(x)[length(x)],
                                 distance <- distance(igraph, x), 
                                 path = paste0(names(x), collapse = '>'))) %>%
    reduce(rbind) 


  map_data <- map_data %>% left_join(stops) 


  names(map_data) <- c("stop_id", "distance", "path", "stop_lat", "stop_lon", "stop_name")

  map_data <- map_data %>%
    mutate(subway_mins = distance%/%60)


  result <- map_data %>% 
    mutate(walking_time = 720*Distance(home$lat, home$lon, stop_lat, stop_lon), 
           total_time = (walking_time + distance),
           walking_mins = walking_time%/%60, 
           total_time_mins = total_time%/%60) 

  # consider n_stops stops near home that lead to the smallest total time
  n_stops <- 1
  start <- result %>% arrange(total_time_mins) %>% head(n_stops)

  itin <- data.frame()
  for (i in 1:n_stops) {
    itin_n <- get_itinerary_directed(igraph, dest, start$stop_id[i], 3, stops)
    # itin_n %>% View
    itin <- rbind(itin, itin_n)
  }
   
  # itin %>% View

  map <- map_an_itinerary(home, start, itin)

  return(map)
}
```


#### show before/after L train from around Metropolitan Ave (BK) to 14 St (NYC)
```{r}
map_with <- get_map_with_walking(home, graph)
map_without <- get_map_with_walking(home, l_graph)
sync(map_with, map_without)

```



#### L train heatmap
##### with L
```{r make-grid}
#### Create a grid of points
map_data <- get.all.shortest.paths(graph, dest,  mode = "out")

map_data <- map_data$res %>% 
  lapply(function(x)data.frame(stop_id = names(x)[length(x)],
                               distance <- distance(graph, x), 
                               path = paste0(names(x), collapse = '>'))) %>%
  reduce(rbind) 

map_data <- map_data %>% left_join(stops) 

names(map_data) <- c("stop_id", "distance", "path", "stop_lat", "stop_lon", "stop_name")

map_data <- map_data %>%
  mutate(subway_mins = distance%/%60)

map_data_lats <- seq(min(map_data$stop_lat), max(map_data$stop_lat), by=0.002)
map_data_lons <- seq(min(map_data$stop_lon), max(map_data$stop_lon), by=0.002)
grid <- expand.grid(map_data_lats, map_data_lons) %>%
 select(lat = Var1, lon = Var2)
```


#### Get subway travel time, walking time, total time
```{r get-times}
map_data$path <- as.character(map_data$path)

total_time <- vector()
walking <- vector()
station <- vector()
path = vector()

for(i in 1:nrow(grid)){
 result<- map_data %>% mutate(walking_time = 720*Distance(grid$lat[i], grid$lon[i], stop_lat, stop_lon), 
                              total_time = (walking_time + distance))
 idx <- which.min(result$total_time)

 total_time[i] <- result[idx, "total_time"]
 walking[i] <- result[idx, "walking_time"]
 station[i] <- result[idx, "stop_name"]
 path[i] <- result[idx, "path"]
 # print(walking[i])
}
grid$time <- total_time
grid$walking <- walking
grid$nearest_stop <- station
grid$old_path <- path
grid <- grid %>% mutate(total_time_mins = time%/%60, walking_mins = walking%/%60, subway_mins = total_time_mins - walking_mins)
grid$log_mins <- log(grid$total_time_mins)


```


```{r make-l-grid}
#### Create a grid of points
l_map_data <- get.all.shortest.paths(l_graph, dest,  mode = "out")

l_map_data <- l_map_data$res %>% 
  lapply(function(x)data.frame(stop_id = names(x)[length(x)],
                               distance <- distance(l_graph, x), 
                               path = paste0(names(x), collapse = '>'))) %>%
  reduce(rbind) 

l_map_data <- l_map_data %>% left_join(stops) 

names(l_map_data) <- c("stop_id", "distance", "path", "stop_lat", "stop_lon", "stop_name")

l_map_data <- l_map_data %>%
  mutate(subway_mins = distance%/%60)

l_grid <- expand.grid(map_data_lats, map_data_lons) %>%
 select(lat = Var1, lon = Var2)
```


##### without L
#### Get subway travel time, walking time, total time
```{r get-l-times}
l_map_data$path <- as.character(l_map_data$path)

total_time <- vector()
walking <- vector()
station <- vector()
path = vector()

for(i in 1:nrow(l_grid)){
 result<- l_map_data %>% mutate(walking_time = 720*Distance(l_grid$lat[i], l_grid$lon[i], stop_lat, stop_lon), 
                              total_time = (walking_time + distance))
 idx <- which.min(result$total_time)

 total_time[i] <- result[idx, "total_time"]
 walking[i] <- result[idx, "walking_time"]
 station[i] <- result[idx, "stop_name"]
 path[i] <- result[idx, "path"]
 # print(walking[i])
}
l_grid$time <- total_time
l_grid$walking <- walking
l_grid$nearest_stop <- station
l_grid$path <- path
l_grid <- l_grid %>% mutate(total_time_mins = time%/%60, walking_mins = walking%/%60, subway_mins = total_time_mins - walking_mins)
l_grid$log_mins <- log(l_grid$total_time_mins)


```


#### Calculate differences in total, subway, and walking times with/without the L
```{r}
time_diffs <- cbind(grid, select(l_grid, path, l_total_time_mins = total_time_mins, l_walking_mins = walking_mins, l_subway_mins = subway_mins)) %>% filter(old_path != path) 

time_diffs <- time_diffs %>%
  mutate(total_diff = l_total_time_mins - total_time_mins,
         subway_diff = l_subway_mins - subway_mins,
         walking_diff = l_walking_mins - walking_mins)
```


### Important formatting
##### Filter for points reachable within 30 mins of walking
```{r}

filtered_grid <- time_diffs %>%
  filter(walking_mins <= 30 | l_walking_mins <= 30) %>%
  mutate(min_interval = cut(total_time_mins, seq(0, max(total_time_mins), by = 5), include.lowest = TRUE),
         bin = cut(total_time_mins, seq(0, max(total_time_mins), by = 5), include.lowest = TRUE, labels = FALSE))

filtered_grid_spdf <- filtered_grid
coordinates(filtered_grid_spdf) <- ~lon + lat
proj4string(filtered_grid_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(filtered_grid_spdf, nyc_neighborhoods)
filtered_grid <- cbind(filtered_grid, matches)
filtered_grid <- filtered_grid %>%
  filter(!is.na(neighborhood))

coords <- filtered_grid[c("lon", "lat")]

gridded(filtered_grid_spdf) <- TRUE

# gray plot
# plot(filtered_grid_spdf, cex = 0.2, col = "grey")
# points(coords, pch =1, col = "black", cex = 0.1)
```


#### Map the difference in total time
```{r}
#Interpolate surface and fix the output
idw <- idw(formula=sqrt(total_diff) ~ 1, locations=filtered_grid_spdf, newdata = filtered_grid_spdf)  # apply idw model for the data

residual_grid <- raster(idw, "var1.pred")
contourr <- rasterToContour(residual_grid)


rr <- mask(residual_grid, nyc_neighborhoods)
crs(rr) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

clrs <- brewer.pal(11, "RdYlGn")

grid_pal <- colorNumeric(
  palette = clrs,
  domain = sqrt(filtered_grid$total_diff),
  reverse = TRUE,
  na.color = "transparent")

m_total <- leaflet(nyc_neighborhoods) %>% 
  addTiles() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  setView(-73.87, 40.75, zoom = 11) %>%
  addLegend("bottomright", pal = grid_pal, values = sqrt(filtered_grid$total_diff),
    title = paste("Change in Total Commute Time", "<br/>", "(areas affected by the L)"),
    labFormat = labelFormat(suffix = " mins", transform = function(x) x * x),
    opacity = 0.5) %>% 
  addRasterImage(rr, color=grid_pal, opacity = 0.5)

```

#### Map the difference in subway time
```{r}
#Interpolate surface and fix the output
idw <- idw(formula=subway_diff ~ 1, locations=filtered_grid_spdf, newdata = filtered_grid_spdf)  # apply idw model for the data

residual_grid <- raster(idw, "var1.pred")
contourr <- rasterToContour(residual_grid)


rr <- mask(residual_grid, nyc_neighborhoods)
crs(rr) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

clrs <- brewer.pal(11, "RdYlGn")

grid_pal <- colorNumeric(
  palette = clrs,
  domain = filtered_grid$subway_diff,
  reverse = TRUE,
  na.color = "transparent")

m_subway <- leaflet(nyc_neighborhoods) %>% 
  addTiles() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  setView(-73.87, 40.75, zoom = 11) %>%
  addLegend("bottomright", pal = grid_pal, values = filtered_grid$subway_diff,
    title = paste("Change in Subway Time", "<br/>", "(areas affected by the L)"),
    labFormat = labelFormat(suffix = " mins"),
    opacity = 0.5) %>% 
  addRasterImage(rr, color=grid_pal, opacity = 0.5)


```

#### Map the difference in walking time
```{r}
#Interpolate surface and fix the output
idw <- idw(formula=walking_diff ~ 1, locations=filtered_grid_spdf, newdata = filtered_grid_spdf)  # apply idw model for the data

residual_grid <- raster(idw, "var1.pred")
contourr <- rasterToContour(residual_grid)


rr <- mask(residual_grid, nyc_neighborhoods)
crs(rr) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

clrs <- brewer.pal(11, "RdYlGn")

grid_pal <- colorNumeric(
  palette = clrs,
  domain = filtered_grid$walking_diff,
  reverse = TRUE,
  na.color = "transparent")

m_walking <- leaflet(nyc_neighborhoods) %>% 
  addTiles() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  setView(-73.87, 40.75, zoom = 11) %>%
  addLegend("bottomright", pal = grid_pal, values = filtered_grid$walking_diff,
    title = paste("Change in Walking Time", "<br/>", "(areas affected by the L)"),
    labFormat = labelFormat(suffix = " mins"),
    opacity = 0.5) %>% 
  addRasterImage(rr, color=grid_pal, opacity = 0.5)

```

```{r}
sync(m_total, m_subway, m_walking)
```